## Study 1: Number agreement in English {#study1}

```{r agreementpreparemodels, echo=FALSE, message=FALSE, warnings=FALSE}
AgreementNullModel <- glm(Agreement~POS, data = AgreementData, family=binomial)
AgreementNullReport <- ReportGLM(AgreementNullModel, AgreementData$Agreement, name="Null Model", intercept.descriptor = "(Adjective)")
AgreementPaperModel <- glm(Agreement~POS*Dialect, data = AgreementData, family=binomial)
AgreementPaperReport <- ReportGLM(AgreementPaperModel, AgreementData$Agreement, name="Quasi-Reported Model", intercept.descriptor = "(Adjective, South, 0)")
AgreementTextModel <- glmer(Agreement~POS*Dialect+(1|Text), data = AgreementData, family=binomial)
AgreementTextReport <- ReportGLM(AgreementTextModel, AgreementData$Agreement, name="Text Model", intercept.descriptor = "(Adjective, South, 0)")
```

### What was reported {#study1reported}

Study 1 concerns the loss of plural agreement marking within the NP during the Middle English period, read as a reduction of redundancy when plural marking disappears on attributive adjectives and determiners. The hypothesis attributes this partly to the influence of Scandinavian languages between the eighth and eleventh centuries, and treats the differing pace at which determiners and adjectives shed their marking as itself possibly contact-related. The study examines determiners (called *quantifiers* throughout, though not all determiners are quantificational) and monosyllabic adjectives across 47 texts totalling 4,366 data points (called *tokens* on p. 551).

From a plot overlaying texts as dots on a map of England (their Figure 1), the authors argue that some texts show stronger loss of inflection on the determiner than on the adjective.^[Including the borders of the Danelaw would have helped here; the visual pattern is in any case not obvious to us.] @WalkdenEa2023 [552] claim that "this difference does not depend on the date of composition of the manuscript, but rather on the dialectal region. The majority of texts which show a difference between quantifiers and adjectives come from areas where Scandinavian contact was intense." They estimate a linear model "with the difference in agreement between adjectives and quantifiers as our dependent variable and region as independent variable." With the Southern dialects as the reference category (least Scandinavian contact), they report a significant effect for the North and East Midlands but not the West [@WalkdenEa2023, 552], where Footnote 8 was omitted:

> The results showed a significant effect of region for the North ($\beta$=0.39, SE=0.15, t=2.6, p=0.01) and the East Midlands ($\beta$=0.2, SE=0.10, t=2.09, p=0.04), but not the West ($\beta$=0.13, SE=0.13, t=1.0, p=0.3). This geographical distribution suggests that contact with Scandinavian speakers played a relevant role in the loss of inflectional morphology, but equally importantly, it shows that concord in quantifiers was more affected than concord with adjectives.

The inferential logic is pure NHST: every result with p<0.05 is called *significant*. We return below to why *significant* is not *meaningful*. First, though, a more fundamental problem with the dependent variable. The paper never states clearly how the "difference in agreement" was measured; it cannot be a direct observation, since it is a *difference* computed per text. The data package resolves it: for each text the authors took (i) the proportion of determiners with plural agreement and (ii) the proportion of adjectives with plural agreement, and (iii) subtracted the second from the first. This per-text difference score is the model's response variable --- the aggregation whose dangers we set out in Section \@ref(models), now realised on actual data.

```{r agreementpreparedifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  .agr <- AgreementAggregateData[order(AgreementAggregateData$Difference),]
  .grandtotals <- .agr$AdjectiveCount+.agr$DeterminerCount
  .cols <- as.integer(.agr$Dialect)
  .max <- max(.grandtotals)
  .min <- min(.grandtotals)
  .maxi <- which(.grandtotals==.max)
  .mini <- which(.grandtotals==.min)
  .countbiasR <- .agr$AdjectiveCount/.agr$DeterminerCount
  .cbmin <- order(.countbiasR)[1]
  .cbmax <- order(.countbiasR)[length(.countbiasR)]
```

Rather than rehearse the objection in the abstract, we plot the resulting data and let it expose itself. Figure \@ref(fig:agreementdifferencescores) shows all `r nrow(.agr)` texts on the x-axis, sorted by difference score, with the score on the y-axis. We have added two pieces of information the paper does not reveal. The *shape* of each dot encodes its dialectal region (see legend); the *size* of each dot is proportional to the total number of observations behind it (determiners plus adjectives), which ranges from `r .grandtotals[.mini][1]` for *`r .agr[.mini,"Text"][1]`* to `r .grandtotals[.maxi]` for *`r .agr[.maxi,"Text"]`*. The vertical bar on each dot encodes a further property: how strongly determiners are overrepresented relative to adjectives in that text.^[A text naturally contains more determiners than adjectives; the point is that the ratio varies wildly between the chosen texts.] The bar scale differs from the y-axis scale and long bars have been clipped to fit, so the printed numbers should be read for the full picture. *`r .agr[.cbmin,"Text"]`* contributes `r .agr[.cbmin,"AdjectiveCount"]` adjectives and `r .agr[.cbmin,"DeterminerCount"]` determiners --- `r round(.agr[.cbmin,"DeterminerCount"]/.agr[.cbmin,"AdjectiveCount"],2)` times more determiners than adjectives --- whereas *`r .agr[.cbmax,"Text"]`* contributes `r .agr[.cbmax,"AdjectiveCount"]` adjectives and `r .agr[.cbmax,"DeterminerCount"]` determiners, the reverse imbalance.

```{r agreementdifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="100%", fig.cap="Difference scores for the texts in the sample; sizes of the dots correspond to the number of observations contributed by the text; vertical bars indicate the overrepresentation of determiners"}

  par(mar = c(4, 2.5, 1, 0))

  plot(NULL, xlim=c(0,48), ylim=c(-0.6,1.3), bty="n", ylab="Difference score", xlab="", xaxt="n")

  # Plot visual helper lines.
  for (i in 1:nrow(.agr)) {
    lines(c(i,i), c(-0.65, 1.0), col="lightgray", lty = 1, lwd = 1)
  }
  
  # Plot 0-line.
  lines(x = c(1, nrow(.agr)), rep( 0.1, 2), lwd = 2, col="lightgray")
  lines(x = c(1, nrow(.agr)), rep( 0  , 2), lwd = 2, col="gray")
  lines(x = c(1, nrow(.agr)), rep(-0.1, 2), lwd = 2, col="lightgray")

  for (i in 1:nrow(.agr)) {
     # This scales the determiner bias so it fits within the plot.
    .biascoord <- .agr[i, 8]+(1/.countbiasR[i])/10
    .biascoord <- ifelse(.biascoord>1,1,.biascoord)
    
    # Draw the total count bar.
    lines(x=c(i,i), y=c(.agr[i, 8], .biascoord), col=.cols[i], lwd=1, lty=1)
    text(x=i, y=.biascoord+0.1,
         col=.cols[i], labels = round(1/.countbiasR[i],2), cex=0.7, srt=90)

    # Then draw the point for the text
    points(i, .agr[i, 8], pch=.cols[i]+14, cex=4*.grandtotals[i]/.max, col=.cols[i])
  }
  axis(1, 1:nrow(.agr), labels=.agr$Text, las=2, cex=0.5, tick = F, cex.axis=0.5, mgp=c(0,0,0))
  legend(0.5, 1.2, col=1:4, legend = levels(.agr$Dialect), pch=15:18, box.col="white", box.lwd=0, pt.cex=1, bg="white")
```

The plot makes the damage concrete. The model behind the reported result treats the dot for the smallest text --- `r .grandtotals[.mini][1]` observations --- as exactly as informative as the dot for the largest, at `r .grandtotals[.maxi]`. A plain linear model on these scores is the wrong tool for aggregated proportions in the first place,^[Models for proportion data exist, but they are harder to handle and interpret [@ZuurEa2009]; in any case the authors report a plain linear model, not a weighted binomial model on proportions.] and the imbalance compounds the problem: were *Innocendium* and *Horses* as large as *Purvey* and *Ayenbyte*, the generalisations about the Southern group would weaken considerably. The difference score is itself an estimate resting on wildly varying sample sizes and adjective-to-determiner ratios, so it carries an inaccuracy the simple linear model is not built to absorb. Nominal significance levels almost certainly overstate the real ones. This is elementary enough that an author, reviewer, or editor should have caught it.^[Whether more data have since been annotated, or the picture has since clarified, is beside the point: @WalkdenEa2023 is published, peer-reviewed research, and it is this paper we examine. It had been downloaded 1,440 times as of 29 March 2025.]

### Alternative models

The unaggregated data are, fortunately, in the data package. The relevant linguistic event --- even in the Middle Ages --- was a single writer producing a single determiner or adjective, marked or unmarked. The right model therefore estimates the probability that plural agreement is overtly realised on a given word.^[The authors themselves follow this strategy in Study 2 (our Section \@ref(study2)) --- though there they introduce a different, equally inadmissible aggregated score as a regressor.] We specified a model implementing the paper's idea while avoiding the difference-score problem: the response is the probability of overt agreement on a word, as a function of its part of speech (adjective or determiner), dialect group, and their interaction. We call it the **Quasi-Reported Model** --- not what *was* reported, but what *should have been*. Table \@ref(tab:agreementpapertable) gives the coefficients and p-values.

```{r agreementpapertable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  kable(AgreementPaperReport$Table, caption=AgreementPaperReport$Report, 
        align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

Granting a cavalier 0.05 threshold for the sake of argument, the estimates suggest that Midlands writers (East and West) drop agreement more readily overall, while Northern writers do not, but that Northern writers do show a relatively strong tendency to avoid agreement specifically on determiners (the POSDeterminer × DialectNorth interaction). That last point is in line with the paper. The overall dialect picture, however, is not: the West Midlands show an even stronger general tendency to avoid agreement than the East Midlands, which runs against the paper's claim. We do not lean on this as a counter-finding --- the West Midlands estimate rests on its own limited data --- but that is precisely the point. The dialect signal is too thin to support an inference in *either* direction; read honestly, it is noise.

To be fair to the dialect predictors, dropping dialect from the specification yields a model that is worse by standard metrics such as R², which indicates the regressors do carry *some* influence. We call this the Null Model (Table \@ref(tab:agreementnulltable)).

```{r agreementnulltable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementNullReport$Table, caption=AgreementNullReport$Report,
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

```{r agreementnullcompare, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  .t <- compare_performance(AgreementNullModel, AgreementPaperModel)[,c(1,9)]
  colnames(.t) <- c("Model", "Tjur's R²")
  .t[,1] <- c("Null Model", "Quasi-Reported Model")
  .t[,2] <- round(.t[,2], 3)
  .t %>%
    kable(caption="Model comparison for the Null Model and the Quasi-Reported Model",
          booktabs = T, linesep = "") %>%
    row_spec(0, bold=TRUE)
```

But the dots plot already warned us there are simply not many data from the North. That DialectNorth fails to reach significance on its own may reflect nothing more than scarce Northern data combined with a genuinely small effect. The plot also shows texts scattered across the difference-score range regardless of dialect, with most clustered near zero; at the negative extreme, just five East Midlands texts account for every negative value. The honest response is to model variation at the document level as a random effect, absorbing differences in the circumstances, register, and individual style of each text. Table \@ref(tab:agreementtexttable) reports the resulting **Text Model**.

```{r agreementtexttable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementTextReport$Table, caption=AgreementTextReport$Report, 
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

The result is decisive: variation *between texts* swamps every other factor, and the model's explanatory power rises sharply --- conditional R²=`r round(unname(r2(AgreementTextModel)[["R2_conditional"]]),3)`, against Tjur's R²=`r round(unname(r2(AgreementPaperModel)[[1]]),3)` for the Quasi-Reported Model.^[R² measures invite the usual caveats and are not always comparable across model types, but a gap this large is not an artefact.] The only generalisation surviving at the nominal threshold is the POSDeterminer × DialectNorth interaction --- Northern writers avoiding agreement on determiners more than on adjectives. And that rests on `r length(which(AgreementData$Dialect=="North"))` data points from `r length(which(AgreementAggregateData$Dialect=="North"))` texts spanning `r max(AgreementData$Year)-min(AgreementData$Year)` years. Where one must work with so little, it is worth asking whether statistics is the right instrument at all.

### Inspection of the models

```{r agreementprepareconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
.confusionPaper <- table(ifelse(predict(AgreementPaperModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionNull <- table(ifelse(predict(AgreementNullModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionText <- table(ifelse(predict(AgreementTextModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
```

Beyond testing, a fitted binomial model also predicts: for each combination of regressors it estimates a probability, and a threshold (conventionally 0.5) turns that into a binary call. We check the predictions on the very `r length(predict(AgreementPaperModel, type="response"))` observations the model was fitted on, where it should perform at its best; new data would only be harder.

The Quasi-Reported Model (Table \@ref(tab:agreementpaperconfusion)) predicts correctly in `r sum(diag(.confusionPaper))` cases and incorrectly in `r sum(.confusionPaper)-sum(diag(.confusionPaper))`, or `r round(sum(diag(.confusionPaper))/sum(.confusionPaper)*100,1)`% --- barely above the base rate of `r round(length(which(AgreementData$Agreement==1))/nrow(AgreementData)*100,1)`% (the accuracy of always guessing the modal outcome). The reason is laid bare by a single fact: the model assigns only `r .p <- predict(AgreementPaperModel, type="response"); length(table(.p))` distinct probabilities across all `r length(predict(AgreementPaperModel, type="response"))` data points. It sorts every observation into one of `r length(table(.p))` bins. A model that cannot tell its data points apart more finely than that is not differentiating; it is rounding.

```{r agreementpaperconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(.confusionPaper,
      caption="Correct and incorrect predictions of the Quasi-Reported Model",
      booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

The Null Model (not tabulated) is weaker still: it always predicts overt agreement and so does not differentiate at all. The Text Model (Table \@ref(tab:agreementtextvonfusion)) does markedly better, correct in `r sum(diag(.confusionText))` cases and wrong in `r sum(.confusionText)-sum(diag(.confusionText))`, or `r round(sum(diag(.confusionText))/sum(.confusionText)*100,1)`%.

```{r agreementtextvonfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(.confusionText,
      caption="Correct and incorrect predictions of the Text Model",
      booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

The same contrast appears if we vary the decision threshold from 0 to 1 rather than fixing it at 0.5 (sometimes worthwhile, as in rare-events regression with base rates near 99% [@KingZeng2001]). Figure \@ref(fig:agreementpapertextcorrects) tracks accuracy across thresholds for both models. The Quasi-Reported Model sits flat at the base rate until a threshold near 0.8, then collapses to its complement --- the signature of a model that is not really discriminating. The Text Model instead peaks near 0.5, as a well-behaved binomial model should, rising above the base rate because it genuinely separates the cases.

```{r agreementpapertextcorrects, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Percent of binary predictions based on the Quasi-Reported Model and the Text Model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementPaperModelCorrects <- unlist(lapply(.ts, .correct))

.p <- predict(AgreementTextModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementTextModelCorrects <- unlist(lapply(.ts, .correct))

par(mfrow=c(1,2))
plot(.AgreementPaperModelCorrects, type="l", lwd = the.lwd, bty="n",
     main="Quasi-Reported Model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
abline(h=0.71, col="darkgray", lty = 3, lwd = the.lwd)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
plot(.AgreementTextModelCorrects, type="l", lwd = the.lwd, bty="n",
     main="Text Model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
abline(h=0.71, col="darkgray", lty = 3, lwd = the.lwd)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
par(mfrow=c(1,1))
```

Precision and recall tell the same story from a third angle. *Precision* is the share of correct hits among all predicted occurrences (true positives over all positives); *recall* is the share of actual occurrences the model catches (true positives over all real events). Tracked across thresholds (Figure \@ref(fig:agreementpapertextprecrec)), the two curves cross --- their balance point --- around 80% for the Quasi-Reported Model and around 90% for the Text Model: once more, the Text Model differentiates better.

```{r agreementpapertextprecrec, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Precision and Recall for the Quasi-Reported Model and the Text Model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementPaperModelPR <- lapply(.ts, .precrec)
.PrecisionPaper <- unlist(.AgreementPaperModelPR)[seq(1,201,2)]
.RecallPaper <- unlist(.AgreementPaperModelPR)[seq(2,202,2)]

.p <- predict(AgreementTextModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementTextModelPR <- lapply(.ts, .precrec)
.PrecisionText <- unlist(.AgreementTextModelPR)[seq(1,201,2)]
.RecallText <- unlist(.AgreementTextModelPR)[seq(2,202,2)]

par(mfrow=c(1,2))
plot(.PrecisionPaper, type="l", lwd = the.lwd, bty="n",
     main="Quasi-Reported Model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col="darkred"
     )
lines(.RecallPaper, col="darkgreen", lwd = the.lwd, lty = 2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c("darkred","darkgreen"), lty = c(1, 2), lwd = the.lwd, bty="n")
plot(.PrecisionText, type="l", lwd = the.lwd, bty="n",
     main="Text Model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col="darkred"
     )
lines(.RecallText, col="darkgreen", lwd = the.lwd, lty = 2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c("darkred","darkgreen"), lty = c(1, 2), lwd = the.lwd, bty="n")
par(mfrow=c(1,1))
```

### Final remarks

The inferences in the paper are unwarranted: once the inadmissible difference-score aggregation is removed, the data of Study 1 do not support the network of hypotheses of Section \@ref(hypotheses). The regressors' influence on the choice between marked and unmarked pre-nominal elements is so slight that only the standard NHST misreading of p-values could register it as relevant, and the dialectal signal is not merely small but internally contradictory --- there is no clean North-East to South-West cline. Once the individual text enters as a random effect, the one factor with even moderate explanatory power is the text itself --- a proxy for style and register --- not dialect. And that random effect is not an optional flourish: omitting it violates the independence-of-observations assumption the model rests on. Study 1 thus corroborates an old philological commonplace: individual texts differ, and deserve to be read individually.

```{r agreementinspectdata, echo=FALSE, message=FALSE, warnings=FALSE}

  tbl.agrdial <- table(AgreementData$Agreement, AgreementData$Dialect)
  tbl.agrpos  <- table(AgreementData$Agreement, AgreementData$POS)
  tbl.posdial <- table(AgreementData$POS, AgreementData$Dialect)

  ptbl.agrdial <- round(prop.table(table(AgreementData$Agreement, AgreementData$Dialect), margin = 2), 2)
  ptbl.agrpos  <- round(prop.table(table(AgreementData$Agreement, AgreementData$POS), margin = 2), 2)
  ptbl.posdial <- round(prop.table(table(AgreementData$POS, AgreementData$Dialect), margin = 2), 2)

  .data <- AgreementData[which(AgreementData$Dialect == "North"), ]
  n.tbl.agrpos  <- table(.data$Agreement, .data$POS)
  n.ptbl.agrpos <- round(prop.table(table(.data$Agreement, .data$POS), margin = 2), 2)
  
  .data <- AgreementData[which(AgreementData$Dialect == "South"), ]
  s.tbl.agrpos  <- table(.data$Agreement, .data$POS)
  s.ptbl.agrpos <- round(prop.table(table(.data$Agreement, .data$POS), margin = 2), 2)
  
  ptbl.diadist <- round(prop.table(table(AgreementData$Dialect))*100, 2)
  tbl.diadist  <- table(AgreementData$Dialect)
  diadist.n    <- nrow(AgreementData)

  .ddata <- AgreementData[-which(AgreementData$Text == "Richard-rolle"),]
  ptbl.diadist.richeroll <- round(prop.table(table(.ddata$Dialect))*100, 2)
  tbl.diadist.richeroll  <- table(.ddata$Dialect)
  diadist.richeroll.n    <- nrow(.ddata)

```

A closing word on the line between exploratory description and inferential statistics. A look at the data does suggest that the North lost pre-nominal agreement more sharply than other regions, and that determiners there retained even less agreement than adjectives. It looks promising --- and we insist on proper specification and evaluation anyway, not to spoil the result but because the whole purpose of inferential statistics is to mark the boundary between (i) describing a sample, (ii) inferring to a population, and (iii) inferring to a causal network of hypotheses. Not every observed difference licenses the further steps; whether it does depends on how much data there is and how large the difference.

Here the differences do not survive correct modelling for a simple reason: there is too little data. Of the $n=`r diadist.n`$ data points, only `r tbl.diadist["North"]` --- about `r ptbl.diadist["North"]`% --- come from the North, while the West Midlands, where the per-text scores are most scattered, contribute by far the most (`r tbl.diadist["WestMidlands"]`, or `r ptbl.diadist["WestMidlands"]`%). Worse, removing a single dominant Northern text, *Richard-rolle*, leaves just `r tbl.diadist.richeroll["North"]` Northern data points from four texts, or `r ptbl.diadist.richeroll["North"]`% of the reduced sample ($n=`r diadist.richeroll.n`$). Every claim about *the North* --- across several centuries --- thus rests on one text plus `r tbl.diadist.richeroll["North"]` further pre-nominal elements. The findings may be *encouraging*; they cannot bear the weight of the inference placed on them. And the difference-score aggregation is exactly what disguises this, collapsing `r diadist.n` data points into `r nrow(AgreementAggregateData)` and granting every text equal evidential weight regardless of size.
