## Study 1: Number agreement in English {#study1}

```{r agreementpreparemodels, echo=FALSE, message=FALSE, warnings=FALSE}
AgreementNullModel <- glm(Agreement~POS, data = AgreementData, family=binomial)
AgreementNullReport <- ReportGLM(AgreementNullModel, AgreementData$Agreement, name="Null Model", intercept.descriptor = "(Adjective)")
AgreementPaperModel <- glm(Agreement~POS*Dialect, data = AgreementData, family=binomial)
AgreementPaperReport <- ReportGLM(AgreementPaperModel, AgreementData$Agreement, name="Quasi-Reported Model", intercept.descriptor = "(Adjective, South, 0)")
AgreementTextModel <- glmer(Agreement~POS*Dialect+(1|Text), data = AgreementData, family=binomial)
AgreementTextReport <- ReportGLM(AgreementTextModel, AgreementData$Agreement, name="Text Model", intercept.descriptor = "(Adjective, South, 0)")
```

### What was reported {#study1reported}

Study 1 concerns the loss of plural agreement marking within the NP during the Middle English period, read as a reduction of redundancy when plural marking disappears on attributive adjectives and determiners. The hypothesis attributes this partly to the influence of Scandinavian languages between the eighth and eleventh centuries, and treats the differing pace at which determiners and adjectives shed their marking as itself possibly contact-related. The study examines determiners (called *quantifiers* throughout, though not all determiners are quantificational) and monosyllabic adjectives across "47 texts" (p. 551) totalling 4,366 data points (called "tokens" on p. 551).^[We feel uncomfortable having to add this footnote, but without it our reanalysis would appear to be inconsistent.
In the published paper, the authors claim 47 texts. A closer look at the published script where data are prepared and analysed shows that they had adjective data for 44 texts and determiner data for 46 texts. To see this, examine `length(levels(as.factor(ex.dataadj$Text)))` and `length(levels(as.factor(ex.dataquant$Text)))` at line 137 of their file `number.R` where data loading and preparation is done.
The source of this discrepancy appears to be that some texts contained no annotated adjectives whatsoever after the authors performed some cleanups not described in the published paper. However, they start with 48 (not 47) labels for the Text variable. We used the 44 texts referenced both in the adjective and the determiner data set in Figure \@ref(fig:agreementdifferencescores) as we consider the difference score to be undefined in the two cases without adjectives.
A difference of two ratios is undefined if one of the ratios is $0\div 0$.
In the models below, we use all 46 texts because random effects models handle levels with zero data points well.]

From a plot overlaying texts as dots on a map of England (their Figure 1), the authors argue that some texts show stronger loss of inflection on the determiner than on the adjective.^[Including the borders of the Danelaw would have helped here. The visual pattern is in any case not obvious to us.] @WalkdenEa2023 [552] claim that "this difference does not depend on the date of composition of the manuscript, but rather on the dialectal region. The majority of texts which show a difference between quantifiers and adjectives come from areas where Scandinavian contact was intense." They estimate a linear model "with the difference in agreement between adjectives and quantifiers as our dependent variable and region as independent variable." With the Southern dialects as the reference category (least Scandinavian contact), they report a significant effect for the North and East Midlands but not the West [@WalkdenEa2023, 552, where Footnote 8 was omitted]:

> The results showed a significant effect of region for the North ($\beta$=0.39, SE=0.15, t=2.6, p=0.01) and the East Midlands ($\beta$=0.2, SE=0.10, t=2.09, p=0.04), but not the West ($\beta$=0.13, SE=0.13, t=1.0, p=0.3). This geographical distribution suggests that contact with Scandinavian speakers played a relevant role in the loss of inflectional morphology, but equally importantly, it shows that concord in quantifiers was more affected than concord with adjectives.

The inferential logic is pure NHST: every result with $p<0.05$ is called *significant*. We return below to why *significant* is not *meaningful*. First, though, there is a more fundamental problem with the dependent variable as the paper never states clearly how the "difference in agreement" was measured. It cannot be a direct observation, since it is called a *difference* computed per text. The data package resolves this. For each text the authors took (i) the proportion of determiners with plural agreement and (ii) the proportion of adjectives with plural agreement, and (iii) subtracted the second from the first. This per-text difference score is the model's response variable. It is exactly the type of aggregation whose dangers we set out in Section \@ref(models), now realised on actual data.

```{r agreementpreparedifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  .agr <- AgreementAggregateData[order(AgreementAggregateData$Difference),]
  .grandtotals <- .agr$AdjectiveCount+.agr$DeterminerCount
  .cols <- as.integer(.agr$Dialect)
  .max <- max(.grandtotals)
  .min <- min(.grandtotals)
  .maxi <- which(.grandtotals==.max)
  .mini <- which(.grandtotals==.min)
  .countbiasR <- .agr$AdjectiveCount/.agr$DeterminerCount
  .cbmin <- order(.countbiasR)[1]
  .cbmax <- order(.countbiasR)[length(.countbiasR)]
```

Rather than rehearse the objection in the abstract, we plot the resulting data and let it expose itself. Figure \@ref(fig:agreementdifferencescores) shows all `r nrow(.agr)` texts on the x-axis, sorted by difference score, with the score on the y-axis. We have added two pieces of information the paper does not reveal. The shape and colour of each dot encode its dialectal region (see legend) and the *size* of each dot is proportional to the total number of observations behind it (determiners plus adjectives), which ranges from `r .grandtotals[.mini][1]` for *`r .agr[.mini,"Text"][1]`* to `r .grandtotals[.maxi]` for *`r .agr[.maxi,"Text"]`*. The vertical bar on each dot encodes a further property, namely how strongly determiners are overrepresented relative to adjectives in that text.^[A text naturally contains more determiners than adjectives; the point is that the ratio varies wildly between the chosen texts. The published paper hides these imbalances and consequently does not explain them.] The bar scale differs from the y-axis scale and long bars have been clipped to fit, so the printed numbers should be read for the full picture. *`r .agr[.cbmin,"Text"]`* contributes `r .agr[.cbmin,"AdjectiveCount"]` adjectives and `r .agr[.cbmin,"DeterminerCount"]` determiners --- `r round(.agr[.cbmin,"DeterminerCount"]/.agr[.cbmin,"AdjectiveCount"],2)` times more determiners than adjectives --- whereas *`r .agr[.cbmax,"Text"]`* contributes `r .agr[.cbmax,"AdjectiveCount"]` adjectives and `r .agr[.cbmax,"DeterminerCount"]` determiners, the reverse imbalance.

```{r agreementdifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="100%", fig.cap="Difference scores for the texts in the sample; sizes of the dots correspond to the number of observations contributed by the text; vertical bars indicate the overrepresentation of determiners"}

  par(mar = c(4, 2.5, 1, 0))

  plot(NULL, xlim=c(0,48), ylim=c(-0.6,1.3), bty="n", ylab="Difference score", xlab="", xaxt="n")

  # Plot visual helper lines.
  for (i in 1:nrow(.agr)) {
    lines(c(i,i), c(-0.65, 1.0), col="lightgray", lty = 1, lwd = 1)
  }
  
  # Plot 0-line.
  lines(x = c(1, nrow(.agr)), rep( 0.1, 2), lwd = 2, col="lightgray")
  lines(x = c(1, nrow(.agr)), rep( 0  , 2), lwd = 2, col="gray")
  lines(x = c(1, nrow(.agr)), rep(-0.1, 2), lwd = 2, col="lightgray")

  for (i in 1:nrow(.agr)) {
     # This scales the determiner bias so it fits within the plot.
    .biascoord <- .agr[i, 8]+(1/.countbiasR[i])/10
    .biascoord <- ifelse(.biascoord>1,1,.biascoord)
    
    # Draw the total count bar.
    lines(x=c(i,i), y=c(.agr[i, 8], .biascoord), col=.cols[i], lwd=1, lty=1)
    text(x=i, y=.biascoord+0.1,
         col=.cols[i], labels = round(1/.countbiasR[i],2), cex=0.7, srt=90)

    # Then draw the point for the text
    points(i, .agr[i, 8], pch=.cols[i]+14, cex=4*.grandtotals[i]/.max, col=.cols[i])
  }
  axis(1, 1:nrow(.agr), labels=.agr$Text, las=2, cex=0.5, tick = F, cex.axis=0.5, mgp=c(0,0,0))
  legend(0.5, 1.2, col=1:4, legend = levels(.agr$Dialect), pch=15:18, box.col="white", box.lwd=0, pt.cex=1, bg="white")
```

The plot makes the damage concrete. The model behind the reported result treats the dot for the smallest text (`r .grandtotals[.mini][1]` observations) as exactly as informative as the dot for the largest (`r .grandtotals[.maxi]` observations). A plain linear model on these scores is the wrong tool for aggregated proportions in the first place,^[Models for proportion data exist, but they are harder to handle and interpret [@ZuurEa2009]. In any case the authors report a plain linear model, not a weighted binomial model on proportions.] and the imbalance compounds the problem. Were *Innocendium* and *Horses* as large as *Purvey* and *Ayenbyte*, the generalisations about the Southern group would weaken considerably. The difference score is itself an estimate resting on wildly varying sample sizes and adjective-to-determiner ratios, so it carries an inaccuracy the simple linear model is not built to absorb. Nominal significance levels almost certainly overstate the real ones. This is elementary enough that an author, reviewer, or editor should have caught it.^[Whether more data have since been annotated, or the picture has since clarified, is beside the point we are making, although we honestly hope that it is the case. @WalkdenEa2023 as examined here is published, peer-reviewed research, and it had been downloaded 1,968 times as of 31 May 2026 according to the publisher's website.]

### Alternative models

The unaggregated data are, fortunately, in the data package. The relevant linguistic event --- even in the Middle Ages --- was a single writer producing a single determiner or adjective, marked or unmarked. The right model therefore estimates the probability that plural agreement is overtly realised on a given word.^[The authors themselves follow this strategy in Study 2 (our Section \@ref(study2)), though there they introduce a different, equally inadmissible aggregated score as a regressor.] We specified a model implementing the paper's idea while avoiding the difference-score problem. In that model the response is the probability of overt agreement on a word, as a function of its part of speech (adjective or determiner), dialect group, and their interaction. We call it the **Quasi-Reported Model**, not what *was* reported, but what *should have been*. Table \@ref(tab:agreementpapertable) gives the coefficients and p-values.

```{r agreementpapertable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  kable(AgreementPaperReport$Table, caption=AgreementPaperReport$Report, 
        align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

Granting a cavalier 0.05 threshold for the sake of argument, the estimates suggest that Midlands writers (East and West) drop agreement more readily overall, while Northern writers do not, but that Northern writers do show a relatively strong tendency to avoid agreement specifically on determiners (the POSDeterminer × DialectNorth interaction). That last point is in line with the paper. The overall dialect picture, however, is not: the West Midlands show an even stronger general tendency to avoid agreement than the East Midlands, which runs against the paper's claim. We do not lean on this as a counter-finding as the West Midlands estimate rests on its own limited data. However, that is precisely the point. The dialect signal is too thin to support an inference in either direction. Read honestly, it is noise.

To be fair to the dialect predictors, dropping dialect from the specification yields a model that is worse by standard metrics such as R², which indicates the regressors do carry some influence. We call this the Null Model (Table \@ref(tab:agreementnulltable)).

```{r agreementnulltable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementNullReport$Table, caption=AgreementNullReport$Report,
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

```{r agreementnullcompare, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  .t <- compare_performance(AgreementNullModel, AgreementPaperModel)[,c(1,9)]
  colnames(.t) <- c("Model", "Tjur's R²")
  .t[,1] <- c("Null Model", "Quasi-Reported Model")
  .t[,2] <- round(.t[,2], 3)
  .t %>%
    kable(caption="Model comparison for the Null Model and the Quasi-Reported Model",
          booktabs = T, linesep = "") %>%
    row_spec(0, bold=TRUE)
```

But the dots plot already warned us there are simply not many data from the North. That DialectNorth fails to reach significance on its own may reflect nothing more than scarce Northern data combined with a genuinely small effect. The plot also shows texts scattered across the difference-score range regardless of dialect, with most clustered near zero; at the negative extreme, just five East Midlands texts account for every negative value. The honest response is to model variation at the document level as a random effect, absorbing differences in the circumstances, register, and individual style of each text. Table \@ref(tab:agreementtexttable) reports the resulting **Text Model**.

```{r agreementtexttable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementTextReport$Table, caption=AgreementTextReport$Report, 
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

The result is decisive: variation *between texts* swamps every other factor, and the model's explanatory power rises sharply. The conditional R²=`r round(unname(r2(AgreementTextModel)[["R2_conditional"]]),3)`, against Tjur's R²=`r round(unname(r2(AgreementPaperModel)[[1]]),3)` for the Quasi-Reported Model.^[Comparison of R² measures is fraught with peril, as we pointed out in Section \@ref(models). However, a gap this large likely is no artefact.] The only generalisation surviving at the nominal threshold is the POSDeterminer × DialectNorth interaction --- Northern writers avoiding agreement on determiners more than on adjectives. And that rests on `r length(which(AgreementData$Dialect=="North"))` data points from `r length(which(AgreementAggregateData$Dialect=="North"))` texts spanning `r max(AgreementData$Year)-min(AgreementData$Year)` years. Where one must work with so little, it is worth asking whether statistics is the right instrument at all.

### Inspection of the models

```{r agreementprepareconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
.confusionPaper <- table(ifelse(predict(AgreementPaperModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionNull <- table(ifelse(predict(AgreementNullModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionText <- table(ifelse(predict(AgreementTextModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
```

Beyond testing, a fitted binomial model also predicts, and the minimal requirement is that it predict its own input data with high accuracy.
If it does, it differentiates at least somewhat between cases with and without agreement as seen in the data (external validity aside).
For each combination of regressors it estimates a probability, and a threshold (conventionally 0.5) turns that into a binary call. We check the predictions on the very `r length(predict(AgreementPaperModel, type="response"))` observations the model was fitted on.
The Quasi-Reported Model (Table \@ref(tab:agreementpaperconfusion)) predicts correctly in `r sum(diag(.confusionPaper))` cases and incorrectly in `r sum(.confusionPaper)-sum(diag(.confusionPaper))`, or `r round(sum(diag(.confusionPaper))/sum(.confusionPaper)*100,1)`% --- barely above the base rate of `r round(length(which(AgreementData$Agreement==1))/nrow(AgreementData)*100,1)`% (the accuracy of always guessing the modal outcome).
It lands in that territory by overpredicting Agreement as the outcome.
The Null Model (not tabulated) is weaker still. It always predicts overt agreement and so does not differentiate at all.
In that regard, the Quasi-Reported Model is slightly better.
However, the Text Model (Table \@ref(tab:agreementtextvonfusion)) does markedly better, correct in `r sum(diag(.confusionText))` cases and wrong in `r sum(.confusionText)-sum(diag(.confusionText))`, or `r round(sum(diag(.confusionText))/sum(.confusionText)*100,1)`% prediction accuracy.

```{r agreementpaperconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
.confusionPaper %>%
  as.data.frame.matrix() %>%
  tibble::rownames_to_column("Predicted") %>%
  kable(caption = "Correct and incorrect predictions of the Quasi-Reported Model",
        booktabs = TRUE, linesep = "") %>%
  add_header_above(c(" " = 1, "Actual" = 2), bold = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, bold = TRUE)
```

```{r agreementtextvonfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
.confusionText %>%
  as.data.frame.matrix() %>%
  tibble::rownames_to_column("Predicted") %>%
  kable(caption = "Correct and incorrect predictions of the Text Model",
        booktabs = TRUE, linesep = "") %>%
  add_header_above(c(" " = 1, "Actual" = 2), bold = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(1, bold = TRUE)
```

The same contrast appears if we vary the decision threshold from 0 to 1 rather than fixing it at 0.5. Figure \@ref(fig:agreementpapertextcorrects) tracks accuracy across thresholds for both models. The Quasi-Reported Model sits flat at the base rate until a threshold near 0.8, then collapses to its complement.
This is the signature of a model that is not really discriminating. The Text Model instead peaks near 0.5, as a well-behaved binomial model should, rising above the base rate because it genuinely separates the cases.
Finally, precision and recall tell the same story from a third angle, see Figure \@ref(fig:agreementpapertextprecrec).

```{r agreementpapertextcorrects, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Percent of binary predictions based on the Quasi-Reported Model and the Text Model at varying thresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementPaperModelCorrects <- unlist(lapply(.ts, .correct))

.p <- predict(AgreementTextModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementTextModelCorrects <- unlist(lapply(.ts, .correct))

par(mfrow=c(1,2))
plot(.AgreementPaperModelCorrects, type="l", lwd = the.lwd, bty="n",
     main="Quasi-Reported Model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
abline(h=0.71, col="darkgray", lty = 3, lwd = the.lwd)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
plot(.AgreementTextModelCorrects, type="l", lwd = the.lwd, bty="n",
     main="Text Model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
abline(h=0.71, col="darkgray", lty = 3, lwd = the.lwd)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
par(mfrow=c(1,1))
```


```{r agreementpapertextprecrec, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Precision and Recall for the Quasi-Reported Model and the Text Model at varying thresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementPaperModelPR <- lapply(.ts, .precrec)
.PrecisionPaper <- unlist(.AgreementPaperModelPR)[seq(1,201,2)]
.RecallPaper <- unlist(.AgreementPaperModelPR)[seq(2,202,2)]

.p <- predict(AgreementTextModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementTextModelPR <- lapply(.ts, .precrec)
.PrecisionText <- unlist(.AgreementTextModelPR)[seq(1,201,2)]
.RecallText <- unlist(.AgreementTextModelPR)[seq(2,202,2)]

par(mfrow=c(1,2))
plot(.PrecisionPaper, type="l", lwd = the.lwd, bty="n",
     main="Quasi-Reported Model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col="darkred"
     )
lines(.RecallPaper, col="darkgreen", lwd = the.lwd, lty = 2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c("darkred","darkgreen"), lty = c(1, 2), lwd = the.lwd, bty="n")
plot(.PrecisionText, type="l", lwd = the.lwd, bty="n",
     main="Text Model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col="darkred"
     )
lines(.RecallText, col="darkgreen", lwd = the.lwd, lty = 2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c("darkred","darkgreen"), lty = c(1, 2), lwd = the.lwd, bty="n")
par(mfrow=c(1,1))
```

### Final remarks

The reanalysis shows that the inferences in the paper are unwarranted.
Once the inadmissible difference-score aggregation is removed, the data of Study 1 do not support the network of hypotheses of Section \@ref(hypotheses). The regressors' influence on the choice between marked and unmarked pre-nominal elements is so slight that only the standard NHST misreading of p-values could register it as relevant, and the dialectal signal is not merely small but internally contradictory. There is no clean North-East to South-West cline. Once the individual text enters as a random effect, the one factor with even moderate explanatory power is the text itself (also absorbing style and register) not dialect.
Most importantly, including such a powerful random effect is not optional.
Omitting it violates the independence-of-observations assumption the model rests on. Study 1 thus corroborates at best the philological insight that individual texts differ and deserve to be read individually.

```{r agreementinspectdata, echo=FALSE, message=FALSE, warnings=FALSE}

  tbl.agrdial <- table(AgreementData$Agreement, AgreementData$Dialect)
  tbl.agrpos  <- table(AgreementData$Agreement, AgreementData$POS)
  tbl.posdial <- table(AgreementData$POS, AgreementData$Dialect)

  ptbl.agrdial <- round(prop.table(table(AgreementData$Agreement, AgreementData$Dialect), margin = 2), 2)
  ptbl.agrpos  <- round(prop.table(table(AgreementData$Agreement, AgreementData$POS), margin = 2), 2)
  ptbl.posdial <- round(prop.table(table(AgreementData$POS, AgreementData$Dialect), margin = 2), 2)

  .data <- AgreementData[which(AgreementData$Dialect == "North"), ]
  n.tbl.agrpos  <- table(.data$Agreement, .data$POS)
  n.ptbl.agrpos <- round(prop.table(table(.data$Agreement, .data$POS), margin = 2), 2)
  
  .data <- AgreementData[which(AgreementData$Dialect == "South"), ]
  s.tbl.agrpos  <- table(.data$Agreement, .data$POS)
  s.ptbl.agrpos <- round(prop.table(table(.data$Agreement, .data$POS), margin = 2), 2)
  
  ptbl.diadist <- round(prop.table(table(AgreementData$Dialect))*100, 2)
  tbl.diadist  <- table(AgreementData$Dialect)
  diadist.n    <- nrow(AgreementData)

  .ddata <- AgreementData[-which(AgreementData$Text == "Richard-rolle"),]
  ptbl.diadist.richeroll <- round(prop.table(table(.ddata$Dialect))*100, 2)
  tbl.diadist.richeroll  <- table(.ddata$Dialect)
  diadist.richeroll.n    <- nrow(.ddata)

```

A closing word on the line between exploratory description and inferential statistics is in order. A look at the data might hint --- in some researcher's eye --- that the North lost pre-nominal agreement more sharply than other regions, and that determiners there retained even less agreement than adjectives. To some, the data might look promising. We insist on proper model specification and evaluation anyway, not to spoil the result but because the whole purpose of statistics is to mark the boundary between (i) describing a sample, (ii) inferring to a population, and (iii) inferring to a causal network of hypotheses. Not every observed difference licenses the further steps. Whether it does depends on how much data there is and how large the difference.

Here the differences do not survive correct modelling at minimum for the reason that there are too few and too biased data. Of the $n=`r diadist.n`$ data points, only `r tbl.diadist["North"]` (about `r ptbl.diadist["North"]`%) come from the North, while the West Midlands, where the per-text scores are most scattered, contribute by far the most (`r tbl.diadist["WestMidlands"]`, or `r ptbl.diadist["WestMidlands"]`%). Worse, removing a single dominant Northern text, *Richard-rolle*, leaves just `r tbl.diadist.richeroll["North"]` Northern data points from four texts, or `r ptbl.diadist.richeroll["North"]`% of the reduced sample ($n=`r diadist.richeroll.n`$). Every claim about *the North* across several centuries thus rests on one text plus `r tbl.diadist.richeroll["North"]` further pre-nominal elements. Researchers are free to find these findings encouraging, but they cannot bear the weight of the inference placed on them in the case at hand. Blunt-force NHST inference and the difference-score aggregation are exactly what disguises this, collapsing `r diadist.n` data points into `r nrow(AgreementAggregateData)` (or `r length(levels(as.factor(AgreementData$Text)))` if you consider division by zero admissible arithmetic) and granting every text equal evidential weight regardless of size.
