## Study 1: Number agreement in English {#study1}

```{r agreementpreparemodels, echo=FALSE, message=FALSE, warnings=FALSE}
AgreementNullModel <- glm(Agreement~POS, data = AgreementData, family=binomial)
AgreementNullReport <- ReportGLM(AgreementNullModel, AgreementData$Agreement, name="Null Model", intercept.descriptor = "(Adjective)")
AgreementPaperModel <- glm(Agreement~POS*Dialect, data = AgreementData, family=binomial)
AgreementPaperReport <- ReportGLM(AgreementPaperModel, AgreementData$Agreement, name="Paper Model", intercept.descriptor = "(Adjective, South, 0)")
AgreementTextModel <- glmer(Agreement~POS*Dialect+(1|Text), data = AgreementData, family=binomial)
AgreementTextReport <- ReportGLM(AgreementTextModel, AgreementData$Agreement, name="Text Model", intercept.descriptor = "(Adjective, South, 0)")
```

### What was reported {#study1reported}

Study 1 deals with the loss of plural agreement marking in the NP during the Middle English period.
It is interpreted as a reduction of redundancy when plural markings are lost on attributive adjectives and determiners.
Per hypothesis, this is partially due to the influence of Scandianavian languages from the eighth until the eleventh century.
Also, the different pace by which determiners and adjectives lose their agreement markings is analysed as possibly related to language contact.
The study looks at determiners (referred to as *quantifiers* throughout, although not all determiners are quantificational) and monosyllabic adjectives in 47 texts totalling 4,366 data points (called *tokens* on p. 551).

Based on a plot overlaying texts as dots on a map of England (their Figure 1), it is argued that some texts show stronger loss of inflection on the determiner than on the adjective.^[We would like to point out that it would have helped to also include the borders of the Danelaw in the plot. Also, the visual pattern does not appear to be very obvious to us.]
@WalkdenEa2023 [552] claim that "this difference does not depend on the date of composition of the manuscript, but rather on the dialectal region. The majority of texts which show a difference between quantifiers and adjectives come from areas where Scandinavian contact was intense."
The parameters of a linear model "with the difference in agreement between adjectives and quantifiers as our dependent variable and region as independent variable" are estimated.
With the Southern dialects as the reference category (because they had the least amount contact with Scandinavian languages), the authors find a *significant effect* for North and the East Midlands but not the West Midlands [@WalkdenEa2023,552], where Footnote 8 was omitted:

> The results showed a significant effect of region for the North ($\beta$=0.39, SE=0.15, t=2.6, p=0.01) and the East Midlands ($\beta$=0.2, SE=0.10, t=2.09, p=0.04), but not the West ($\beta$=0.13, SE=0.13, t=1.0, p=0.3). This geographical distribution suggests that contact with Scandinavian speakers played a relevant role in the loss of inflectional morphology, but equally importantly, it shows that concord in quantifiers was more affected than concord with adjectives.

Apparently, the inferential logic is pure NHST.
Any result with p<0.05 is called *significant*.
We will return to these results in the next section and show why *significant* does not mean *meaningful* or *substantive*.
However, there is a more fundamental problem to be discussed first.
The paper is very unclear with respect to the measurement of the difference in agreement between adjectives and quantifiers as used as the dependent variable.
It cannot be a direct observation because it is described as a *difference* of some kind, and it appears to be a score calculated for entire texts, probably as some kind of proportion.
Luckily, the data package reveals what was actually computed.
Per text, the authors calculated (i) the proportion of determiners with plural agreement, (ii) the proportion of adjectives with plural agreement, and (iii) subtracted the former from the latter.
The linear model uses this difference score as the dependent variable (or response variable).

```{r agreementpreparedifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  .agr <- AgreementAggregateData[order(AgreementAggregateData$Difference),]
  .grandtotals <- .agr$AdjectiveCount+.agr$DeterminerCount
  .cols <- as.integer(.agr$Dialect)
  .max <- max(.grandtotals)
  .min <- min(.grandtotals)
  .maxi <- which(.grandtotals==.max)
  .mini <- which(.grandtotals==.min)
  .countbiasR <- .agr$AdjectiveCount/.agr$DeterminerCount
  .cbmin <- order(.countbiasR)[1]
  .cbmax <- order(.countbiasR)[length(.countbiasR)]
```

Instead of discussing problems with such an approach from a theoretical point of view, we plot the resulting data to expose them by example.
Figure \@ref(fig:agreementdifferencescores) shows all `r nrow(.agr)` texts (x-axis), sorted by their difference score from lowest to highest, with the difference scores on the y-axis.
However, we added more information that is not revealed in the paper itself.
The shape of the dots corresponds to the dialectal regions (see legend), and their size corresponds linearly to the total number of observations (summed counts of determiners and adjectives), a number that ranges from `r .grandtotals[.mini][1]` for *`r .agr[.mini,"Text"][1]`* to `r .grandtotals[.maxi]` for *`r .agr[.maxi,"Text"]`*.
The vertical bar for each dot encodes one further remarkable numerical property of the texts and their scores:
it shows how differently determiners are overrepresented in the texts.
It is unclear whether the counts provided by the authors are the total numbers of adjectives and determiners in the texts or whether the number represents how many were annotated.
In any case, some texts contribute far less adjectives relative to the number of determiners than others:^[It is not surprising that a text contains more determiners than adjectives. The question is rather why the ratio differs so wildly between the chosen texts.]
Note that the scale of the bars is not the same as the scale of the y-axis and that long lines have been shortened to fit into the plot.
Consequently, the numbers should be inspected for a full picture.
*`r .agr[.cbmin,"Text"]`* contributes `r .agr[.cbmin,"AdjectiveCount"]` adjectives and `r .agr[.cbmin,"DeterminerCount"]` determiners, which means it has `r round(.agr[.cbmin,"DeterminerCount"]/.agr[.cbmin,"AdjectiveCount"],2)` times more determiners than adjectives.
On the other hand, *`r .agr[.cbmax,"Text"]`* contributes `r .agr[.cbmax,"AdjectiveCount"]` adjectives and `r .agr[.cbmax,"DeterminerCount"]` determiners, which means it has `r round(.agr[.cbmax,"DeterminerCount"]/.agr[.cbmax,"AdjectiveCount"],2)` times more adjectives than determiners.

```{r agreementdifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="100%", fig.cap="Difference scores for the texts in the sample; sizes of the dots correspond to the number of observations contributed by the text; vertical bars indicate the overrepresentation of determiners"}

  par(mar = c(4, 2.5, 1, 0))

  plot(NULL, xlim=c(0,48), ylim=c(-0.6,1.3), bty="n", ylab="Difference score", xlab="", xaxt="n")

  # Plot visual helper lines.
  for (i in 1:nrow(.agr)) {
    lines(c(i,i), c(-0.65, 1.0), col="lightgray", lty = 1, lwd = 1)
  }
  
  # Plot 0-line.
  lines(x = c(1, nrow(.agr)), rep( 0.1, 2), lwd = 2, col="lightgray")
  lines(x = c(1, nrow(.agr)), rep( 0  , 2), lwd = 2, col="gray")
  lines(x = c(1, nrow(.agr)), rep(-0.1, 2), lwd = 2, col="lightgray")

  for (i in 1:nrow(.agr)) {
     # This scales the determiner bias so it fits within the plot.
    .biascoord <- .agr[i, 8]+(1/.countbiasR[i])/10
    .biascoord <- ifelse(.biascoord>1,1,.biascoord)
    
    # Draw the total count bar.
    lines(x=c(i,i), y=c(.agr[i, 8], .biascoord), col=.cols[i], lwd=1, lty=1)
    text(x=i, y=.biascoord+0.1,
         col=.cols[i], labels = round(1/.countbiasR[i],2), cex=0.7, srt=90)

    # Then draw the point for the text
    points(i, .agr[i, 8], pch=.cols[i]+14, cex=4*.grandtotals[i]/.max, col=.cols[i])
  }
  axis(1, 1:nrow(.agr), labels=.agr$Text, las=2, cex=0.5, tick = F, cex.axis=0.5, mgp=c(0,0,0))
  legend(0.5, 1.2, col=1:4, legend = levels(.agr$Dialect), pch=15:18, box.col="white", box.lwd=0, pt.cex=1, bg="white")
```

Using such aggregated data to estimate the parameters of a linear model is detrimental.
The unweighted model treats each text as contributing the same amount of information, which simply and obviously is not the case.
The established purpose of such models in empirical sciences is to model the distribution of data points, not their aggregations [see @GelmanHill2006 if in doubt].^[There are models for proportion data and other aggregated data. However, they are much more difficult to handle and---above all---interpret [@ZuurEa2009]. Most crucially, however, the authors report a plain linear model and *not* a weighted binomial model on proportion data.]
Thus, the model from the paper makes the assumption that a text with `r .grandtotals[.mini][1]` data points (the observed mininum) and a text with `r .grandtotals[.maxi]` data points (the observed maximum) can be treated as contributing the same amount of evidence.
This is ill-advised for all types of models, even much simpler ones, as a fictitious example shows.
Assume we had the task of calculating the average length of a newly discovered Italic language, Faloscan.
There are two Faloscan inscriptions found, one containing two short sentences averaging 3.5 words per sentence because it contains two sentences, one 3 words long, the other one 4 words long.
The other one is a longer legal text comprising 25 sentences with an average length of 8.3 words per sentence.
A phylogeneticist includes Faloscan in their database of average sentence lengths, calculating an average length of $(3.5+8.3)\div 2=`r round((8.3+3.5)/2, 2)`$ words per sentence.
Almost anyone would criticise this as being incorrect, as the shorter text should receive a much lower weight in the calculation of the average.^[The correct calculation of the weighted mean in this case is $(0.075\cdot 3.5)+(0.925\cdot 8.3)=`r round((0.075*3.5)+(0.925*8.3), 2)`$ words per sentence.]

However, @WalkdenEa2023 do not merely calculate a simple average, they estimate parameters of a much more complex model based on an even more extremely unbalanced aggregated metric.
For example, if *Innocendium* and *Horses* were texts as long as (or samples from texts as large as) *Purvey* and *Ayenbyte*, the generalisations about the Southern dialect group would be weakened considerably.^[We would like to stress that it is irrelevant whether more data has been annotated in the meantime and whether the picture has become clearer, changed, etc. @WalkdenEa2023 is published research that has undergone peer review, and it is this specific paper we are looking at from a methodological point of view. It has been downloaded 1,440 times as of 29 March 2025.]
This is so elementary, that at least one of the authors, the reviewers, or the editors should have noticed it.

When we specify a model, we assume all kinds of sources of influence on variance in the regressor **per data point**, which are the actual measurements.
The model is specified to relate the systematic (explained) influences (variance in the explanatory variables) to the variance in the response variable, leaving the unexplained variance as residuals.
However, in the case at hand the difference score is itself an estimate based on wildly varying sample sizes and wildly varying balance between the numbers of adjectives and determiners, which means that it comes with a kind of variance---or, plainly, inaccuracy---that the simple linear model and its interpretation are not made for (in the mathematical sense).
This also means that actual significance levels very likely do not correspond to nominal significance levels, and at least using a more conservative significance level than 0.05 would be recommended.
In the next section we report an alternative model of the data from @WalkdenEa2023, and we show what remains of the results advertised in the paper under more accurate modelling.


### Alternative models

Luckily, the original unaggregated data are available in the data package.
Now, even in the middle ages, the relevant linguistic event was the actual production of a single instance of a determiner or an adjective by a writer.
Therefore, the best model in the case at hand is one that estimates the probability of speakers overtly marking plural agreement in a concrete sentence or not.^[Inconsistency, the authors follow such a strategy in Study 2, see our Section \@ref(study2).
However, they use another incorrectly aggregated score as a regressor in that study.]
We specified a model that implements the idea from the paper but which avoids the problems with the unbalanced difference scores.
It is summarised in Table \@ref(tab:agreementpapertable).
The response is the probability of the realisation of overt agreement on a word as a function of is part-of-speech (adjective or determiner), the dialect group, and an interaction between the two.
We call it the Paper Model even if it is not what *was* reported in the paper but what *should have been* reported in the paper.
See Table \@ref(tab:agreementpapertable) for a report of the coefficients and the associated p-values as found by the estimator.

```{r agreementpapertable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  kable(AgreementPaperReport$Table, caption=AgreementPaperReport$Report, 
        align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

The parameter estimates indicate that (assuming a cavalier NHST significance level of 0.05 for the sake of the argument) that writers from the Midlands (East and West) have a stronger tendency to drop agreement in general, but that writer from the North do not (see the negative coefficients and the low p-values).
However, writers from the Northern dialect seem to have a relatively strong tendency to avoid agreement specifically on determiners (see the interaction between POSDeterminer and DialectNorth).
The last point is in line with the claims in the paper.
However, the West midlands even have the stronger tendency than the East Midlands to avoid agreement overall (including on determiners).
This runs contrary to the claim in @WalkdenEa2023.

In all fairness, if we drop the influence of dialect from the model specification, we arrive at a model that---according to standard metrics like R²---is actually worse, which might indicate that the regressors do indeed have some influence on the response.
We call it the Null Model, see Table \@ref(tab:agreementnulltable).

```{r agreementnulltable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementNullReport$Table, caption=AgreementNullReport$Report,
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

```{r agreementnullcompare, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  .t <- compare_performance(AgreementNullModel, AgreementPaperModel)[,c(1,3,9)]
  colnames(.t) <- c("Model", "AIC", "Tjur's R²")
  .t[,1] <- c("Null Model", "Paper Model")
  .t[,c(2,3)] <- round(.t[,c(2,3)], 3)
  .t %>%
    kable(caption="Model comparison for the Null Model and the Paper Model",
          booktabs = T, linesep = "") %>%
    row_spec(0, bold=TRUE)
```

With all this, we should keep in mind that the plot in the previous section shows that there simply are not that many data from the North.
The fact that DialectNorth by itself does not reach the nominal significance level in the GLMs could also be due to the fact that there simply are not enough data points from the North and that the effect itself is, at the same time, not verty large.
Also, the plot shows that individual texts and their dialect group are all over the place in terms of ther difference score, and most texts have a difference score very close to 0 anyway.
At the negative end, for example, there are merely five texts from the East Midlands that account for all the negative values.
Hence, it seems appropriate to model the variance at the document level as a random effect, accounting for the different circumstances under which the documents were written, register differences, differences in individual styles, etc.
The resulting model is reported in Table \@ref(tab:agreementtexttable).

```{r agreementtexttable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementTextReport$Table, caption=AgreementTextReport$Report, 
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

Immediately, we see that the variation between the texts swamps all other explanatory factors while the overall explanatory power of the model goes up considerably, as measured in the conditional R²=`r round(unname(r2(AgreementTextModel)[["R2_conditional"]]),3)` compared to Tjur's R²=`r round(unname(r2(AgreementPaperModel)[[1]]),3)` from the Paper Model.^[While R² measures have been criticised and cannot always be compared between different types of models and types of R², this difference is so considerable that denying it is futile.]
The only relevant generalisation that survives at the nominal significance level is the interaction between POSDeterminer and DialectNorth, i.e., writers from the North avoid agreement on determiners more strongly than on adjectives.
It should be kept in mind that this generalisation is based on `r length(which(AgreementData$Dialect=="North"))` data points from `r length(which(AgreementAggregateData$Dialect=="North"))` texts spanning a period of `r max(AgreementData$Year)-min(AgreementData$Year)` years.
Regardless of nominal significance, a careful interpretation seems wise.
While it is sometimes necessary to work with such tiny amounts of data in historical philology, it is questionable whether statistics are the right tool.
We proceed to a further inspection of the models.

### Inspection of the models

```{r agreementprepareconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
.confusionPaper <- table(ifelse(predict(AgreementPaperModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionNull <- table(ifelse(predict(AgreementNullModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionText <- table(ifelse(predict(AgreementTextModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
```

A (Generalised) Linear (Mixed) Model can be used to make predictions and thus go beyond any testing framework, be it NHST or any more correct inferential framework.
In the case of Binomial models, it can estimate the probability with which the event (in this case overt realisation of a subject pronoun) occurs.
If it estimates a probability greater than 0.5 for some combination of regressors, it predicts the occurrence of the event, otherwise it predicts its non-occurrence.
This probability threshold of 0.5 is successful if the model is approximately correct, but under some extreme conditions, it can be better to adapt it to another value between 0 and 1.
We now check in tabular form what binary predictions the model makes on the 3,773 observations it was estimated on with a threshold of 0.5.
Note that it should perform particularly well on those data.
New data would always be harder to predict correctly for any statistical or machine-learning model.

Table \@ref(tab:agreementpaperconfusion) summarises the predictive accuracy of the Paper Model (which is the GLM that implements the idea from the paper in a mathematically more sound way).
The columns correspond to the actual cases of overt an non-overt agreement in the data, the rows tabulate the decisions based on the model estimates against those.
The predictions are correct in `r sum(diag(.confusionPaper))` cases and incorrect in `r sum(.confusionPaper)-sum(diag(.confusionPaper))` cases, i.e., `r round(sum(diag(.confusionPaper))/sum(.confusionPaper)*100,1)`%.^[R² coefficients correspond with the prediction accuracy, and the two convey a similar notion of the adequacy of the model given the data.]
This is slightly better than the base rate (proportion of the modal category, i.e., the percentage of cases in which agreement is overtly realised) of `r round(length(which(AgreementData$Agreement==1))/nrow(AgreementData)*100,1)`%.^[The base rate is relevant inasmuch as one can usually make above 50% correct predictions simply by always guessing the modal---most frequent---outcome. In the long run, the correct guesses then equal the base rate.]
However, it should be noted that the model predicts only `r .p <- predict(AgreementPaperModel, type="response"); length(table(.p))` numerically different predictions for the `r length(predict(AgreementPaperModel, type="response"))` data points.
In other words, the model puts all data points in one of `r length(table(.p))` bins.
That is not a sign of a model that differentiates well.

```{r agreementpaperconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(.confusionPaper,
      caption="Correct and incorrect predictions of the Paper Model",
      booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

The Null Model (not shown in tabular form) with only part-of-speech as a predictor is considerably weaker inasmuch as it always predicts overt agreement and hence does not differentiate at all.
However, the Text Model (Table \@ref(tab:agreementtextvonfusion)), which is the one that incorporates a random effect for text, does much better than anything reported in the paper.
Its predictions are correct in `r sum(diag(.confusionText))` cases and incorrect in `r sum(.confusionText)-sum(diag(.confusionText))` cases, i.e., `r round(sum(diag(.confusionText))/sum(.confusionText)*100,1)`%.

```{r agreementtextvonfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(.confusionText,
      caption="Correct and incorrect predictions of the Text Model",
      booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

As mentioned above, sometimes models make biased probability estimates, and an adapted threshold (for example, lower than 0.5) for binary decisions leads to better results.
This is often the case in rare events regression [@KingZeng2001], which is used when the base base rate is even higher (for example, around 99%).
We can only expect `r length(table(.p))` steps in this curve as the model only predicts so many different probabilities.
Adjusting the threshold at which we make the binary decision incrementally from 0 to 1, we arrive at Figure \@ref(fig:agreementpapertextcorrects) for the Paper Model (left panel) and the Text Model (right panel).
While it is also not a model that differentiates extremely well, we see that the Text Model is at least able to differentiate beyond the mere prediction of the modal category (thus always landing at the base rate).
The peak is close to 0.5, as expected for regular binomial models.
The Paper Model essentially predicts at the level of the base rate all the way until a threshold of roughly 0.8 is reached, at which point the prediction accuracy drops dramatically to 100 minus the base rate.
The curve has `r length(table(.p))` steps as expected.

```{r agreementpapertextcorrects, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Percent of binary predictions based on the Paper Model and the Text Model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementPaperModelCorrects <- unlist(lapply(.ts, .correct))

.p <- predict(AgreementTextModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementTextModelCorrects <- unlist(lapply(.ts, .correct))

par(mfrow=c(1,2))
plot(.AgreementPaperModelCorrects, type="l", lwd = the.lwd, bty="n",
     main="Paper Model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
abline(h=0.71, col="darkgray", lty = 3, lwd = the.lwd)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
plot(.AgreementTextModelCorrects, type="l", lwd = the.lwd, bty="n",
     main="Text Model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
abline(h=0.71, col="darkgray", lty = 3, lwd = the.lwd)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
par(mfrow=c(1,1))
```

Finally, we can approach the accuracy of the model from two other perspectives.
The *Precision* is the proportion of correct predictions of the event occurring (here, the subject being overtly realised) among all predictions of this event.
In other words, it is the number of true positives divided by all positives (true or false).
The *Recall*, on the other hand, is the proportion of correct predictions of the event occurring among all the cases in which the event actually occurred.
Put differently, it is the number of true positives divided by the number of events that have occurred.
These two scores, evaluated at different thresholds, can be used to assess the quality of the model and adjust the threshold.
The balanced point (where Precision and Recall reach their highest values and the curves cross) is around 80% for the Paper Model and around 90% for the Text Model, which is another indicator that the Text Model differentiates much better.

```{r agreementpapertextprecrec, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Precision and Recall for the Paper Model and the Text Model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementPaperModelPR <- lapply(.ts, .precrec)
.PrecisionPaper <- unlist(.AgreementPaperModelPR)[seq(1,201,2)]
.RecallPaper <- unlist(.AgreementPaperModelPR)[seq(2,202,2)]

.p <- predict(AgreementTextModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementTextModelPR <- lapply(.ts, .precrec)
.PrecisionText <- unlist(.AgreementTextModelPR)[seq(1,201,2)]
.RecallText <- unlist(.AgreementTextModelPR)[seq(2,202,2)]

par(mfrow=c(1,2))
plot(.PrecisionPaper, type="l", lwd = the.lwd, bty="n",
     main="Paper Model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col="darkred"
     )
lines(.RecallPaper, col="darkgreen", lwd = the.lwd, lty = 2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c("darkred","darkgreen"), lty = c(1, 2), lwd = the.lwd, bty="n")
plot(.PrecisionText, type="l", lwd = the.lwd, bty="n",
     main="Text Model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col="darkred"
     )
lines(.RecallText, col="darkgreen", lwd = the.lwd, lty = 2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c("darkred","darkgreen"), lty = c(1, 2), lwd = the.lwd, bty="n")
par(mfrow=c(1,1))
```

### Final remarks

What does all this mean?
It means that the inferences in the paper are unwarranted, and that the data in Study 1 do not provide evidence in favour of the network of hypotheses explicitly or implicitly introduced (see Section \@ref(hypotheses)), especially if the model is reformulated without the inadmissible aggregation in the form of the difference scores.
The influences of the regressors on the choice between overtly marked and unmarked pre-nominal elements are so minor that only typical NHST misinterpretation of p-values could wrongly detect them as having a relevant influence.
What is more, the dialectal region has minor and above all contradictory pseudo-influence inasmuch as there is no clear North-East to South-West continuum.
If the individual text is added as a random effect, we see that the single influence with at least moderate explanatory power is the individual text (and thus likely a combination of style and register), not the dialectal region.
It should be noticed that including the text as a random effect is not even an optional decision as *not* including it violates assumptions underlying the statistical model, in this case predominantly the *independence of events*.
To summarise, the data from Study 1 corroborate an old wisdom from traditional philology, videlicet that individual texts differ and deserve careful individual examination in order to be decoded and understood properly.

```{r agreementinspectdata, echo=FALSE, message=FALSE, warnings=FALSE}

  tbl.agrdial <- table(AgreementData$Agreement, AgreementData$Dialect)
  tbl.agrpos  <- table(AgreementData$Agreement, AgreementData$POS)
  tbl.posdial <- table(AgreementData$POS, AgreementData$Dialect)

  ptbl.agrdial <- round(prop.table(table(AgreementData$Agreement, AgreementData$Dialect), margin = 2), 2)
  ptbl.agrpos  <- round(prop.table(table(AgreementData$Agreement, AgreementData$POS), margin = 2), 2)
  ptbl.posdial <- round(prop.table(table(AgreementData$POS, AgreementData$Dialect), margin = 2), 2)

  .data <- AgreementData[which(AgreementData$Dialect == "North"), ]
  n.tbl.agrpos  <- table(.data$Agreement, .data$POS)
  n.ptbl.agrpos <- round(prop.table(table(.data$Agreement, .data$POS), margin = 2), 2)
  
  .data <- AgreementData[which(AgreementData$Dialect == "South"), ]
  s.tbl.agrpos  <- table(.data$Agreement, .data$POS)
  s.ptbl.agrpos <- round(prop.table(table(.data$Agreement, .data$POS), margin = 2), 2)
  
  ptbl.diadist <- round(prop.table(table(AgreementData$Dialect))*100, 2)
  tbl.diadist  <- table(AgreementData$Dialect)
  diadist.n    <- nrow(AgreementData)

  .ddata <- AgreementData[-which(AgreementData$Text == "Richard-rolle"),]
  ptbl.diadist.richeroll <- round(prop.table(table(.ddata$Dialect))*100, 2)
  tbl.diadist.richeroll  <- table(.ddata$Dialect)
  diadist.richeroll.n    <- nrow(.ddata)

```

A final word regarding the difference between exploratory data analysis and inferential statistics is in order.
We acknowledge that a look at the data might suggest that in the North there is a more pronounced loss of prenominal agreement markers within the NP than in the other regions, and that determiners show proportionally even less agreement markers than adjectives in the North.
It looks promising.
So, why do we ruin the party and insist on a proper model specification and model evaluation instead of celebrating the fact that the authors found the differences they were looking for?
That is because we consider it vital to maintain the difference between (i) descriptive observations in samples and (ii) inferences about the whole population and especially (iii) inferences about complex networks of hypotheses about what causes the observed differences.
This is exactly why we have got inferential statistics in the first place: not any observed difference allows us to make far-reaching inferences.
Whether an inference is admissible depends on the amount of data and the magnitude of the observed difference.

In the case at hand, the reason why the observable differences in the sample do not hold up under correct modelling and do not allow any meaningful inferences is the amount of available data.
There are $n=`r diadist.n`$ data points.
Of these data points, only `r tbl.diadist["North"]` are from the North, which is roughly `r ptbl.diadist["North"]`% of the entire sample.
The West Midlands (where the per-text difference scores are all over the place) contribute the largest amount of data by far (`r tbl.diadist["WestMidlands"]` data points or `r ptbl.diadist["WestMidlands"]`%).
If we remove the one dominant text from the North (*Richard-rolle*), there are only `r tbl.diadist.richeroll["North"]` data points from the North coming from four texts left, which corresponds to `r  ptbl.diadist.richeroll["North"]`% of the reduced data set ($n=`r diadist.richeroll.n`$).
In essence, the authors are basing a huge part our inference (all statements about *the North*) about a period spanning several centuries on one text and `r tbl.diadist.richeroll["North"]` additional occurrences of prenominal elements.
Hence, while the results might be called *encouraging*, they are not sufficient to base any heavy inferences on.
What makes it worse is that the aggregation (the difference score) reported in the paper makes things look much more promising than they are by reducing $`r diadist.n`$ data points to $`r nrow(AgreementAggregateData)`$ aggregated data points, treating all texts as contributing the same amount of evidence.
