## Study 2: Null subjects in Spanish {#study2}

```{r subjectpreparemodels, echo=FALSE, message=FALSE, warnings=FALSE}
SubjectPaperModel <- glmer(Subject~ORScoreZ*YearZ+Macroregion+(1|Text), data = SubjectData, family="binomial")
SubjectPaperReport <- ReportGLM(SubjectPaperModel, SubjectData$Subject, name="Paper Model", intercept.descriptor = "(0, 0, Spain)")
SubjectNullModel <- glmer(Subject~1+(1|Text), data = SubjectData, family="binomial")
SubjectNullReport <- ReportGLM(SubjectNullModel, SubjectData$Subject, name="Text-Only Model", intercept.descriptor = "(1)")
SubjectCountryGenreModel <- glmer(Subject~Country+Genre+(1|Text), data = SubjectData, family="binomial")
SubjectCountryGenreReport <- ReportGLM(SubjectCountryGenreModel, SubjectData$Subject, name="Country-Genre Model", intercept.descriptor = "(Spain, Nonliterary)")
SubjectFullCountryModel <- glmer(Subject~ORScoreZ*YearZ+Country+Genre+(1|Text), data = SubjectData, family="binomial")
SubjectFullCountryReport <- ReportGLM(SubjectFullCountryModel, SubjectData$Subject, name="Full-Country Model", intercept.descriptor = "(0, 0, Spain, DOC)")
SubjectFullNoGenreModel <- glmer(Subject~ORScoreZ*YearZ+Country+(1|Text), data = SubjectData, family="binomial")
SubjectFullNoGenreReport <- ReportGLM(SubjectFullNoGenreModel, SubjectData$Subject, name="Full-No-Genre Model", intercept.descriptor = "(0, 0, Spain)")
```

### What was reported

In Study 2, the authors set out to model the influence of orality, time, and region (Spain versus Latin American countries) on the tendency to use overt subjects in Spanish despite the overall very strong tendency towards subject drop in Spanish.
Orality is measured as the ORScore, which appears to be an aggregate relative frequency of certain grammatical features in the respective texts.
It is thus a derived construct and not a direct observation.
Given the fact that it is composed of proportions (relative frequencies) the warnings raised in the previous section regarding aggregate scores apply here as well (albeit to a regressor and not to the response).
Table \@ref(tab:subjectpapertable) shows the report of the model fit from the paper, where *Subject* is 1 for an observation with an overt subject, *ORScoreZ* is the z-transformed orality score, *YearZ* is the z-transformed year the respective document was dated at, and *Macroregion* is Spain or Non-Spain.
The random effect *Text* has 37 levels for 37 documents.
Henceforth, we shall call this the Paper Model.

```{r subjectpapertable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(SubjectPaperReport$Table, caption=SubjectPaperReport$Report, 
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

We confirm the correctness of our reconstruction of the model reported in the paper by looking at the quote from @WalkdenEa2023 [558] (Footnote 14 omitted) and verifying that the provided numbers match those from our reconstruction of the model.
Highlighting was added by us.
Obviously, we have managed to set up our model exactly as the authors did.^[After double and triple checking, we conclude that $p<0.03$ for ORScoreZ:YearZ as reported in the paper is a minor lapse.]

> However, this trend also holds for Spain, which contradicts our sociolinguistic predictions.
> Since the orality effect discussed in section 4.3 may be at play in the bar charts, it is necessary to run a mixed-effects model to confirm any relationships across countries and centuries.
> A mixed model using glmer from the lme4 package (Bates et al. 2015) in R (R Core Team 2021) was run to find any such underlying patterns.
> The model had pronoun realisation (**3,773 total tokens**) as the binary-dependent variable; macroregion (Spain vs. Non-Spain), orality and year (both z-scored) as fixed effects; and document ID as a random effect.
> Country was originally included as a fixed effect but had too many levels and was replaced with macroregion. The interaction between year and orality was significant (**β=−0.44**, SE=0,21, **p\<0.03**), where the negative coefficient means that for every year that passes, the effect of orality on overtness is lessened.
> Thus, there is an underlying diachronic effect, but as we can see from Figure 3, it is a small effect.
> Macro-region was also found to be just significant (**β=0.63**, SE=0.32, **p\<0.05**), favouring overtness in every country but Spain.
> This result explains the apparent rise we saw for Spain in Figure 3 as just a product of orality: the later Spanish texts score higher in ORSCORE.

A very overt problem that *any* reviewer should have spotted is that in this section of the paper, p-values are given as *less than* (\<), wheras in the report of Study 1, they were given as *equal to* (=), see our Section \@ref(study1reported).
Clearly, the values for both studies are rounded to two decimal points, but why this discrepancy?
Also, the phrase *the orality effect discussed in section 4.3 may be at play in the bar charts* is extremely odd as the orality effect was at play in the language community (*if* it was), not in a plot in a paper from 2023.
Also, *it is necessary to run a mixed-effects model to confirm any relationships across countries and centuries* is incorrect as such a necessity does not exist.
There is never a theoretical necessity to use a random effect.
see @Schaefer2020a for a brief introduction and @GelmanHill2006 [245–247] for a summary of the inconsistencies in such assumptions about when to use random effects.

Most importantly, brute-force NHST inference is used once again:
less than 0.05 leads to a positive inference, greater than 0.05 leads to a negative or no inference.
This reaches almost bizarre levels when `r SubjectPaperReport$Table[4,5]` for the Macroregion coefficient is taken as *just significant*.
However, the authors then explain (p. 558) that "*[t]his result explains the apparent rise we saw for Spain in Figure 3 as just a product of orality: the later Spanish texts score higher in ORSCORE.*"
What they are saying is that their model is misspecified!
If there is considerable covariation between Country or Macroregion and ORSCORE and you know about it, it is detrimental to leave it unmodelled.
An interaction or a similar solution in the mixed-model paradigm is in order.
Also, it is completely unclear why Macroregion was included in the model and not Country (see Section \@ref(study2alt)).
This discussion is obsolete, however, as in Study 2, the model is even less informative than in Study 1, as we will show presently.

### Alternative models {#study2alt}

We begin our reanalysis by comparing the Paper Model to a Text-Only Model (not reported in tabular form), which is a model that only has a fixed intercept and a random intercept for the text.
In other words, we allow the intercept to vary between groups of data points defined by the texts they occur in, and no other explanatory variables are included.
`r SubjectNullReport$Report`
The AIC of the Paper Model is lower.
The marginal R² of the Text-Only Model is of course 0, as there are no first-level fixed effects, but even the conditional R² of the Paper Model is minimally higher than that of the Text-Only Model.
All in all, it does not appear even from this simple evaluation as if the orality score, the macroregion, and the date of the document explain much of the variance in the data, as clearly indicated by the R² measures and the prediction accuracy.

A second alternative model we would like to entertain is the Country-Genre Model, see Table \@ref(tab:subjectcountrygenretable).
It drops the alleged main explanatory regressor (orality), the secondary one (year), and it specifies the country and the genre as fixed effects instead.
Contrary to the claim in the paper, the standard maximum likelihood estimator converges without any additional tweaks, and the Country regressor does not have too many levels at all.^[Even if it did indeed have too many levels for reliable fixed-effect estimation, a random effect (nesting the document ID) could and should have been used, see @GelmanHill2006; @Schaefer2020a.]

```{r subjectcountrygenretable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(SubjectCountryGenreReport$Table, caption=SubjectCountryGenreReport$Report,
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

```{r subjectcountrygenrecompare, warning=F, echo=F}
  .t <- compare_performance(SubjectCountryGenreModel, 
                          SubjectPaperModel)[,c(1,3,9,10)]
  colnames(.t) <- c("Model", "AIC", "Conditional R²", "Marginal R²")
  .t[,1] <- c("Country-Genre Model", "Paper Model")
  .t[,c(2,3,4)] <- round(.t[,c(2,3,4)], 3)
  .t %>%
    kable(caption="Model comparison for the Null Model and the Paper Model",
          booktabs = T, linesep = "") %>%
    row_spec(0, bold=TRUE)
```

We see in Table \@ref(tab:subjectcountrygenrecompare) that the Country-Genre Model is also slightly worse than the Paper Model.
It is unclear, however, why the authors claim that the country regressor was impossible to add for technical reasons.
The estimator has no trouble estimating the coefficients, they just don't reach the nominal NHST significance level of 0.05.

We proceed to estimating the parameters of a more complex model.
The Full-Country Model reported in Table \@ref(tab:subjectfullcountrytable) is the Paper Model with orality and year as regressors but with country instead of macroregion and additionally genre as fixed effects.

```{r subjectfullcountrytable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(SubjectFullCountryReport$Table, 
      caption=SubjectFullCountryReport$Report, align="llrrrc",
      booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

```{r subjectfullcountrycompare, warning=F, echo=F}
  .t <- compare_performance(SubjectFullCountryModel, 
                          SubjectPaperModel)[,c(1,3,9,10)]
  colnames(.t) <- c("Model", "AIC", "Conditional R²", "Marginal R²")
  .t[,1] <- c("Full-Country Model", "Paper Model")
  .t[,c(2,3,4)] <- round(.t[,c(2,3,4)], 3)
  .t %>%
    kable(caption="Model comparison for the Full-Country Model and the Paper Model",
          booktabs = T, linesep = "") %>%
    row_spec(0, bold=TRUE)
```

The Full-Country Model performs slightly better than the Paper Model in terms of the R², as shown in Table \@ref(tab:subjectfullcountrycompare).
It should, however, have become clear by now that the slightly varying conditional R² coefficients are merely artefacts due to the shifting marginal R² coefficients.
Additionally, we observe that country has only one level (Bolivia) that is sufficiently distinct from the intercept value (Spain) to reach the nominal significance level.
Concluding from the data presented in the paper that "overtness [is favoured] in every country but Spain" is thus not warranted, even under an NHST umbrella.
If we drop genre, the picture changes again with the Full-No-Genre Model, see Tables \@ref(tab:subjectfullnogenretable) and \@ref(tab:subjectfullnogenrecompare).

```{r subjectfullnogenretable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(SubjectFullNoGenreReport$Table, caption=SubjectFullNoGenreReport$Report,
      align="llrrrc", booktabs = T, linesep = "",) %>%
  row_spec(0, bold=TRUE)
```

```{r subjectfullnogenrecompare, warning=F, echo=F}
  .t <- compare_performance(SubjectFullNoGenreModel, 
                          SubjectPaperModel)[,c(1,3,9,10)]
  colnames(.t) <- c("Model", "AIC", "Conditional R²", "Marginal R²")
  .t[,1] <- c("Full-No-Genre Model", "Paper Model")
  .t[,c(2,3,4)] <- round(.t[,c(2,3,4)], 3)
  .t %>%
    kable(caption="Model comparison for the Full-No-Genre Model and the Paper Model",
          booktabs = T, linesep = "") %>%
    row_spec(0, bold=TRUE)
```

We see that model specification is not an exact art.
Dropping genre as a regressor leads to Columbia also reaching the nominal significance level.
Based on NHST logic, it is now Spain, the Dominican Republic, and Panamá versus Bolivia and Columbia.
Regarding the R², the various model specifications do not  seem to change by any considerable margin.
In other words, it does not seem to matter which regressors we include, the results are always middling at an almost exactly identical level.
In the next section, we show why, focussing on the Paper Model as one example of the many meaningless models one can specify for the data.

### Inspection of the models

```{r subjectpaperconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
SubjectPaperPredictNull <- length(which(predict(SubjectPaperModel, type="response")<=0.5))
SubjectPaperPredictOne <- length(which(predict(SubjectPaperModel, type="response")>=0.5))
SubjectPaperBaseline <- as.numeric(predict(SubjectPaperModel, type="response")>0.5)
```

As explained in the discussion of Study 1, we check the probability estimates made by the Paper Model for the data used to estimate its parameters.
First, we examine the correct binary predictions using a 0.5 probability threshold.
The Model predicts a null subject in `r SubjectPaperPredictNull` cases and a pronominal subject in `r SubjectPaperPredictOne` cases.
In other words, it predicts that there should never be an overt subject.
This is due to the obviously uninformative regressors and the high base rate.
The base rate (no overt subject) is at `r round(mean(SubjectPaperBaseline==SubjectData$Subject)*100,2)`%.
To see whether the Paper Model could make better binary predictions with a lower probability threshold for binary decisions, we plot the predictions of the Paper Model and the Text-Only Model next to each other as violin plots in Figure \@ref(fig:subjectpaperviolins).

```{r subjectpaperviolins, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Probabilities estimated by the Paper Model and the Text-Only Model; only for data used to estimate the model parameters and grouped by their actual outcome"}
par(mfrow=c(1,2))
vioplot(predict(SubjectPaperModel, type="response")~SubjectData$Subject, col = "lemonchiffon", colMed = "white", colMed2 = "grey 75", xlab = "Paper Model", ylab="Estimated probability", frame.plot = F, ylim = c(0, 0.275))
vioplot(predict(SubjectNullModel, type="response")~SubjectData$Subject, col = "lemonchiffon", colMed = "white", colMed2 = "grey 75", xlab = "Text-Only Model", ylab="", frame.plot = F, ylim = c(0, 0.275))
par(mfrow=c(1,1))
```

Even for the cases with overt subjects in the observed data, the model estimates a maximum probability of under 0.25.
Most strikingly, at least judging from the distribution of the predictions the Paper Model estimates almost exactly the same probabilities as the Text-Only Model.
In other words, orality, time, and country (or macroregion) are by and large irrelevant.
Reaching the significance level is almost exclusively an effect of the relatively simple model structure, minute differences correlated with the regressors, and the relatively large sample size.
Alls points of criticism against NHST apply.
The left panel of Figure \@ref(fig:subjectpapercorrects) shows the percentage of correct binary predictions of the model at different thresholds.
The percentage rises quickly until the point where 100% minus the base rate percentage has been reached, which is indicative of a model that does nothing but predict the modal category, reaching the base rate in the best case. 
In the right panel of Figure \@ref(fig:subjectpapercorrects), we plot the Precision and Recall curves for the Paper Model.
Since the model does not predict any probabilities above `r round(max(.p), 2)`, both precision and recall cannot be higher than 0 past this threshold.

```{r subjectpapercorrects, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Left: Percent of binary predictions based on the Paper Model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome; Right: Precision and Recall for the Paper Model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}

par(mfrow = c(1,2))

.p <- predict(SubjectPaperModel, type="response")
.d <- SubjectData$Subject
.ts <- seq(0, 1, 0.01)

# Contect capture.
.correct <- function(t) Correct(.p, .d, t)

.SubjectPaperModelCorrects <- unlist(lapply(.ts, .correct))

plot(.SubjectPaperModelCorrects, type="l", lwd = the.lwd, bty="n",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
abline(h = 0.89, lwd = the.lwd, lty = 3, col = "darkgray")
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))

# PrecRec

# Context capture.
.precrec <- function(t) PrecisionRecall(.p, .d, t)
  
.SubjectPaperModelPR <- lapply(.ts, .precrec)
.Precision <- unlist(.SubjectPaperModelPR)[seq(1,201,2)]
.Recall <- unlist(.SubjectPaperModelPR)[seq(2,202,2)]

plot(.Precision, type="l", lwd = the.lwd, bty="n",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col="darkred"
     )
lines(.Recall, col="darkgreen", lwd = the.lwd, lty = 2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("topright", legend=c("Precision", "Recall"), col=c("darkred", "darkgreen"), lty = c(1, 2), lwd = the.lwd, bty="n")
```

```{r subjectpapernullcompare, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="80%", fig.cap="Comparing the predictions of the Paper Model and the Text-Only Model"}
plot(predict(SubjectNullModel, type="response"), predict(SubjectPaperModel, type="response"),
     pch=20, bty ="n",
     xlim = c(0,0.3), ylim = c(0,0.3),
     col = "black",
     xlab = "Probabilities estimated under the Text-Only Model",
     ylab = "Probabilities estimated under the Paper Model"
)
```

Finally, and most revealing perhaps, we plot the probabilities predicted by the Paper Model against those predicted by the Text-Only Model in Figure \@ref(fig:subjectpapernullcompare) in the style of a correlation plot.
The plot shows which probability the two models estimate for each of the 3,773 observed data points, each data point being plotted as a dot.
If the two models predicted exactly the same probabilities, the dots should be aligned along a straight line.
Indeed, we see dots mostly aligned along a straight line with negligible deviations.
The models contain virtually the same information, and the mean absolute difference between the predictions by the two models is `r round(mean(abs(predict(SubjectNullModel, type="response")-predict(SubjectPaperModel, type="response"))),4)`, the maximal difference is `r round(max(abs(predict(SubjectNullModel, type="response")-predict(SubjectPaperModel, type="response"))),4)`.

Furthermore and most crucially, we see only 37 dots, not 3,773.
This is because the model makes exactly the same predictions for each data point from the individual texts, and the dots are therefore printed on top of each other.
The model demonstrates (weakly) that individual medieval Spanish texts differ from each other, nothing more.
It does not at all "confirm any relationships across countries and centuries" (see above).
These problems are not at all hard to pinpoint with full access to the data, and it is astonishing that the analysis went on to be submitted in paper form.
We whole-heartedly support publishing negative findings, and hence the data themselves deserve publishing.
However, negative findings should be clearly published as such, and not as corroborations of a massive and unclear network of hypotheses, especially given problems of underdetermination discussed in Section \@ref(findingout).

Study 3 reported in @WalkdenEa2023 is highly preliminary, and we therefore do not discuss it.
The additive models used in the analysis have their own set of problems, and a discussion would take up too much space in the present paper.
In the last section, we will discuss what all this means for substantive hypotheses put forward in @WalkdenEa2023 and the field in general.
