## Study 1: Number agreement in English {#study1}

```{r agreementpreparemodels, echo=FALSE, message=FALSE, warnings=FALSE}
AgreementNullModel <- glm(Agreement~POS, data = AgreementData, family=binomial)
AgreementNullReport <- ReportGLM(AgreementNullModel, AgreementData$Agreement, name="Null model", intercept.descriptor = "(Adjective)")
AgreementPaperModel <- glm(Agreement~POS*Dialect, data = AgreementData, family=binomial)
AgreementPaperReport <- ReportGLM(AgreementPaperModel, AgreementData$Agreement, name="Paper model", intercept.descriptor = "(Adjective, South, 0)")
AgreementTextModel <- glmer(Agreement~POS*Dialect+(1|Text), data = AgreementData, family=binomial)
AgreementTextReport <- ReportGLM(AgreementTextModel, AgreementData$Agreement, name="Text model", intercept.descriptor = "(Adjective, South, 0)")
```

### What was reported

Study 1 deals with the loss of plural agreement marking in the NP during the Middle English period.
It is interpreted as a reduction of redundancy when plural markings are lost on attributive adjectives and determiners.
Per hypothesis, this is partially due to the influence of Scandianavian languages from the eighth until the eleventh century.
Also, the different pace by which determiners and adjectives lose their agreement markings is analysed as possibly related to language contact.
The study looks at determiners (referred to as *quantifiers* throughout, although not all determiners are quantificational) and monosyllabic adjectives in 47 texts totalling 4,366 data points (called *tokens* on p. 551).

Based on a plot overlaying texts as dots on a map of England (their Figure 1), it is argued that some texts show stronger loss of inflection on the determiner than on the adjective.^[We would like to point out that it would have helped to also include the borders of the Danelaw in the plot. Also, the visual pattern does not appear to be very obvious.]
@WalkdenEa2023 [552] claim that "this difference does not depend on the date of composition of the manuscript, but rather on the dialectal region. The majority of texts which show a difference between quantifiers and adjectives come from areas where Scandinavian contact was intense."
The parameters of a linear model "with the difference in agreement between adjectives and quantifiers as our dependent variable and region as independent variable" are estimated.
With the Southern dialects as the reference category (because they had the least amount contact with Scandinavian languages), the authors find a *significant effect* for North and the East Midlands but not the West Midlands [@WalkdenEa2023,552], where Footnote 8 was ommitted:

> The results showed a significant effect of region for the North ($\beta$=0.39, SE=0.15, t=2.6, p=0.01) and the East Midlands ($\beta$=0.2, SE=0.10, t=2.09, p=0.04), but not the West ($\beta$=0.13, SE=0.13, t=1.0, p=0.3). This geographical distribution suggests that contact with Scandinavian speakers played a relevant role in the loss of inflectional morphology, but equally importantly, it shows that concord in quantifiers was more affected than concord with adjectives.

We will return to these results in the next section.
However, the paper is slightly blurry with respect to the measurement of the difference in agreement between adjectives and quantifiers as used as the dependent variable.
It cannot be a direct observation because it is described as a difference of some kind, and it appears to be a score calculated for entire texts, probably as some kind of proportion.
Luckily, the data package reveals what was actually computed.
Per text, the authors calculated (i) the proportion of determiners with plural agreement, (ii) the proportion of adjectives with plural agreement, and (iii) subtracted the former from the latter.
The linear model uses this difference score as the dependent variable.

```{r agreementpreparedifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
  .agr <- AgreementAggregateData[order(AgreementAggregateData$Difference),]
  .grandtotals <- .agr$AdjectiveCount+.agr$DeterminerCount
  .cols <- as.integer(.agr$Dialect)
  .max <- max(.grandtotals)
  .min <- min(.grandtotals)
  .maxi <- which(.grandtotals==.max)
  .mini <- which(.grandtotals==.min)
  .countbiasR <- .agr$AdjectiveCount/.agr$DeterminerCount
  .cbmin <- order(.countbiasR)[1]
  .cbmax <- order(.countbiasR)[length(.countbiasR)]
```

Instead of discussing problems with such an approach from a theoretical point of view, we plot the resulting data to expose them by example.
The plot shows all `r nrow(.agr)` texts (x-axis), sorted by their difference score from lowest to highest, with the difference scores plottet on the y-axis.
However, we added more information that is not revealed in the paper itself.
The shape of the dots corresponds to the dialectal regions (see legend), and their size corresponds linearly to the total number of observations (summed counts of determiners and adjectives), a number that ranges from `r .grandtotals[.mini][1]` for *`r .agr[.mini,"Text"][1]`* to `r .grandtotals[.maxi]` for *`r .agr[.maxi,"Text"]`*.
The vertical bar for each dot encodes one further remarkable numerical property of the texts and their scores:
it shows how differently determiners are overrepresented in the texts.
It is not absolutely clear wheter the counts provided by the authors are the total numbers of adjectives and determiners in the texts or whether the number represents how many were annotated.
In any case, some texts contribute far less adjectives relative to the number of determiners than others:^[It is not surprising that a text contains more determiners than adjectives. The question is rather why the ratio differs so wildly between texts.]
Note that the scale of the bars is not the same as the scale of the y-axis and that long lines have been shortened to fit into the plot.
Consequently, the numbers should be inspected.
*`r .agr[.cbmin,"Text"]`* contributes `r .agr[.cbmin,"AdjectiveCount"]` adjectives and `r .agr[.cbmin,"DeterminerCount"]` determiners, which means it has `r round(.agr[.cbmin,"DeterminerCount"]/.agr[.cbmin,"AdjectiveCount"],2)` times more determiners than adjectives.
On the other hand, *`r .agr[.cbmax,"Text"]`* contributes `r .agr[.cbmax,"AdjectiveCount"]` adjectives and `r .agr[.cbmax,"DeterminerCount"]` determiners, which means it has `r round(.agr[.cbmax,"DeterminerCount"]/.agr[.cbmax,"AdjectiveCount"],2)` times more adjectives than determiners.

```{r agreementdifferencescores, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', out.width="100%", fig.cap="Difference score for the texts in the sample; sizes of the dots correspond to the number of observations contributed by the text; vertical bars indicate the overrepresentation of determiners"}
  plot(NULL, xlim=c(0,48), ylim=c(-0.6,1.3), bty="n", ylab="Difference score", xlab="", xaxt="n")

  # Plot visual helper lines.
  for (i in 1:nrow(.agr)) {
    lines(c(i,i), c(-0.65, 1.0), col="lightgray")
  }
  
  for (i in 1:nrow(.agr)) {
     # This scales the determiner bias so it fits within the plot.
    .biascoord <- .agr[i, 8]+(1/.countbiasR[i])/10
    .biascoord <- ifelse(.biascoord>1,1,.biascoord)
    
    # Draw the total count bar.
    lines(x=c(i,i), y=c(.agr[i, 8], .biascoord), col=.cols[i], lwd=1, lty=1)
    text(x=i, y=.biascoord+0.1,
         col=.cols[i], labels = round(1/.countbiasR[i],2), cex=0.7, srt=90)

    # Then draw the point for the text
    points(i, .agr[i, 8], pch=.cols[i]+14, cex=4*.grandtotals[i]/.max, col=.cols[i])
  }
  axis(1, 1:nrow(.agr), labels=.agr$Text, las=2, cex=0.5, tick = F, cex.axis=0.5, mgp=c(0,0,0))
  legend("bottomright", col=1:4, legend = levels(.agr$Dialect), pch=15:18, box.col="white", box.lwd=0, pt.cex=1.5, bg="white")
```

Using such aggregated data to estimate the parameters of a linear model is unwise.
The unweighted model treats each text as contributing the same information, and its main purpose in empirical sciences is to model the distribution of data points, not aggregations.^[There are models for proportion data and other aggregated data. However, they are much more difficult to handle and---above all---interpret (**REF**). Still, the authors report a plain linear model.]
Thus, the model from the paper makes the assumption that a text with `r .grandtotals[.mini][1]` data points (the observed mininum) and a text with `r .grandtotals[.maxi]` data points (the observed maximum) can be treated as contributing the same amount of evidence.
This is ill-advised for all types of models, even much simpler ones like the ungrouped average of a population, which is a standard example:
Assume we had the task of calculating the average length of a rare kind of newts.
Furthermore, assume their true population average length is 10 cm.
Because our newts are so rare and difficult to find, instead of going out and measuring a random sample of those rare newts, we look for existing data, finding a report where field researchers measured two newts they found in a pond averaging 7 cm in length and another report from a larger lake where researchers measured the average length of a population of 40 newts to be 10.5 cm.
Since the original raw data tables (individual measurements) are not included in the old reports, we average the averages, correctly finding that (7+10.5)/2 is `r (7+10.5)/2`.
While correct, this is certainly not a good estimator of the average length of our newts.
It is severly skewed because of the unequal sizes of the samples, and the two newts from the pond receive much too high a weight.
Similarly, if *Innocendium* and *Horses* were just as long a text (or as large a sample from a text) as *Purvey* and *Ayenbyte*, all generalisations about the Southern dialect group would disappear.^[We would like to stress that it is irrelevant whether more data has been annotated in the meantime and whether the picture has become clearer, changed, etc. @WalkdenEa2023 is published research that has undergone peer review, and it is this specific paper we are looking at from a methodological point of view.]

In other words, when we specify a model we assume that all kinds of sources influence the variance in the data points, which are assumed to be simple measurements.
The model is specified to relate the systematic (explained) influences  (variance in the explanatory variables) to the variance in the response variable, leaving the unexplained variance as residuals.
However, in this case the difference score is itself an estimate based on wildly varying sample sizes and varying balance between the numbers of adjectives and determiners, which means that it comes with a kind of variance---or rather: inaccuracy---that the simple linear model and its interpretation do not expect (in a technical sense).
This also means that actual significance levels very likely do not correspond to nominal significance levels, and using a more rigid significance level than 0.05 would be recommended.
The next section shows an alternative model and what remains of the results advertised in the paper after a more accurate modelling.

### Alternative models

The example given above about the biologists and the length of newts assumed that the original measurements (lengths of individual newts) were lost.
This is not the case here, and the original data points are available in the data package.
Even in the middle ages, the relevant event was the actual production of a single instance of a determiner or an adjective by a writer.
Therefore, the best model for the case at hand is one that estimates the probability of them marking plural agreement or not.^[As a minor inconsistency, the authors follow such a strategy in Study 2, see below.]
A model that comes close to the idea in the paper is the following one.
It models the probability of the realisation of agreement on a word as a function of ist part-of-speech (adjective or determiner), the dialect group, and an interaction between the two.

```{r agreementpapertable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementPaperReport$Table, caption=AgreementPaperReport$Report, align="llrrrc")
```

The GLM estimates indicate that (assuming a cavalier significance level of 0.05 for the sake of the argument) that writers from the Midlands (East and West) have a stronger tendency to drop agreement in general, but that writer from the North do not (see the negative coefficients and the low p vaues).
However, writers from the Northern dialect seem to have a relatively strong tendency to avoid agreement specifically on determiners (see the interaction between POSDeterminer and DialectNorth).
The last point is in line with the claims in the paper.
However, the West midlands even have the stronger tendency than the East Midlands to avoid agreement overall (including on determiners).
This runs contrary to the claim in @WalkdenEa2023.

If we drop the influence of dialect from the model specification, we arrive at a model that---according to standard metrics like R²---is actually worse.

```{r agreementnulltable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementNullReport$Table, caption=AgreementNullReport$Report, align="llrrrc")
```

```{r agreementnullcompare, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(compare_performance(AgreementNullModel, AgreementPaperModel)[,c(1,3,9,10)], caption="Model comparison for the Null model")
```

With all this, we should keep in mind that the plot in the previous section shows that there simply are not that much data from the North.
The fact that the level DialectNorth by itself does not reach the nominal significance level in the GLMs could also be due to the fact that there simply are not enough data points from the North.
Also, the plot shows that individual texts and their dialect group are all over the place in terms of ther difference score, and most texts have a difference score very close to 0 anyway.
At the negative end, for example, there are merely five texts from the East Midlands that account for all the negative values.
It seems appropriate to model the variance at the document level as a random effect, accounting for the different circumstances under which the documents were written, register differences, differences in individual styles, etc.

```{r agreementtexttable, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(AgreementTextReport$Table, caption=AgreementTextReport$Report, align="llrrrc")
```

We see immediately that the variation between the texts swamps all other explanatory factors while the overall explanatory power of the model goes up, at least measured in the conditional R²=`r round(unname(r2(AgreementTextModel)[["R2_conditional"]]),2)`.
The only relevant generalisation that survives at the nominal significance level is the interaction between POSDeterminer and DialectNorth, i.e., writers from the North avoid agreement on determiners more strongly than on adjectives.
It should be kept in mind that this generalisation is based on `r length(which(AgreementData$Dialect=="North"))` data points from `r length(which(AgreementAggregateData$Dialect=="North"))` texts spanning a period of `r max(AgreementData$Year)-min(AgreementData$Year)` years.
Regardless of nominal significance, a careful interpretation seems wise.
We proceed to a further inspection of the models presented in the next section.

### Interpretation of the models

```{r agreementprepareconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
.confusionPaper <- table(ifelse(predict(AgreementPaperModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionNull <- table(ifelse(predict(AgreementNullModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
.confusionText <- table(ifelse(predict(AgreementTextModel, type="response")<0.5,
                                "No Agreement", "Agreement"),
                         ifelse(AgreementData$Agreement==0, "No Agreement", "Agreement"))
```

A (Generalised) Linear (Mixed) Model can be used to make predictions.
In the case of Binomial models, it can estimate the probability with which the event (in this case overt realisation of a subject pronoun) occurs.
If it estimates a probability greater than 0.5 for some combination of regressors, it is the most successful (and customary) strategy to predict the occurrence of the event.^[The strategy is successful only if the model is approximately correct.]
If it estimates a probability lower than 0.5, one should bet against the event and predict its non-occurrence.
We now check in tabular form what binary predictions the model makes on the 3,773 observations it was estimated on.
Note that it should perform particularly well on those data because we use the model to predict the values it was estimated on.
New data would always be harder to predict correctly for any statistical or machine-learning model.
The following table summarises the predictive accuracy of what we shall call the *Paper model* (which is the GLM that implements the idea from the paper in a mathematically more sound way).
The columns correspond to the actual cases of overt an non-overt agreement in the data, the rows tabulate the decisions based on the model estimates against those.
The predictions are correct in `r sum(diag(.confusionPaper))` cases and incorrect in `r sum(.confusionPaper)-sum(diag(.confusionPaper))` cases, i.e., `r round(sum(diag(.confusionPaper))/sum(.confusionPaper)*100,1)` %.^[R² coefficients correspond with the prediction accuracy, and the two convey a similar notion of the adequacy of the model given the data.]
This is slightly better than the base rate (proportion of the modal category, i.e., the percentage of cases in which agreement is overtly realised) of `r round(length(which(AgreementData$Agreement==1))/nrow(AgreementData)*100,1)` %.^[The base rate is relevant inasmuch as one can usually make above 50 % correct predictions simply by always guessing the modal---most frequent---outcome. In the long run, the correct guesses then equal the base rate.]

```{r agreementpaperconfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(.confusionPaper, caption="Correct and incorrect predictions of the Paper model")
```

The *Null model* with only part-of-speech as a predictor is considerably weaker inasmuch as it always predicts overt agreement and hence does not differentiate at all.
However, the *Text model*, which is the one that incorporates a random effect for text, does much better than what was suggested in the paper.
Its predictions are correct in `r sum(diag(.confusionText))` cases and incorrect in `r sum(.confusionText)-sum(diag(.confusionText))` cases, i.e., `r round(sum(diag(.confusionText))/sum(.confusionText)*100,1)` %.

```{r agreementtextvonfusion, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
kable(.confusionText, caption="Correct and incorrect predictions of the Text model")
```

If we look at the numerical probabilities estimated by the Paper model and the Text model for the observed data, we see that the model that accounts for variance between texts (and thus, implicitly, authors), differentiates much better.
In the violin plots, the left shape summarises the distribution of probabilities estimated by the model for cases where no overt agreement was observed in the data.
The right shape summarises the distribution of probabilities for cases where agreement was overtly realised in the data.
Obviously, the Paper model is not even a good model of the data used to estimate its parameters, which means that it is very likely that it does not generalise well to other data.
Put differently, we see clear indication that the external validity of the model is low.
The Text model fares much better in this respect, but only by acknowledging that the main influencing factor is the text (who wrote the text, under which circumstances, etc.).

```{r agreementpapertextviolins, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', fig.cap="Probabilities estimated by the Paper model and the Text model; only for data used to estimate the model parameters and grouped by their actual outcome"}
  par(mfrow=c(1,2))
  vioplot(predict(AgreementPaperModel, type="response")~AgreementData$Agreement, colMed = "white", colMed2 = "grey 75", xlab = "Paper model", ylab="Estimated probability")
  vioplot(predict(AgreementTextModel, type="response")~AgreementData$Agreement, colMed = "white", colMed2 = "grey 75", xlab = "Model with text as random effect", ylab="")
  par(mfrow=c(1,1))
```

However, sometimes models make biased probability estimates, and an adapted threshold (for example, lower than 0.5) for binary decisions leads to better results.
This can be the case with rare events regression (**King&Tseng**), which is the case with even much higher base rates (around 99%).
Adjusting the threshold at which we make the binary decision incrementally from 0 to 1, we arrive at the following plot for the Paper and the Text model.
While it is not a model that differentiates very well, we see that the Text model is at least able to differentiate beyond a prediction of the base rate.
The peak is close to 0.5, as expected for normal binomial models.

```{r agreementpapertextcorrects, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', fig.cap="Percent of binary predictions based on the Paper model and the Text model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementPaperModelCorrects <- unlist(lapply(.ts, .correct))

.p <- predict(AgreementTextModel, type="response")
.d <- AgreementData$Agreement
.ts <- seq(0, 1, 0.01)
.correct <- function(t) Correct(.p, .d, t)
.AgreementTextModelCorrects <- unlist(lapply(.ts, .correct))

par(mfrow=c(1,2))
plot(.AgreementPaperModelCorrects, type="l", lwd=2, bty="n",
     main="Paper model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
plot(.AgreementTextModelCorrects, type="l", lwd=2, bty="n",
     main="Text model",
     xlab="Probability threshold for prediction",
     ylab="Percent of correct predictions"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     )
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
par(mfrow=c(1,1))
```

Finally, we can approach the accuracy of the model from two other perspectives.
The *Precision* is the proportion of correct predictions of the event occurring (here, the subject being overtly realised) among all predictions of this event.
In other words, it is the number of true positives divided by all positives (true or false).
The *Recall*, on the other hand, is the proportion of correct predictions of the event occurring among all the cases in which the event actually occurred.
Put differently, it is the number of true positives divided by the number of events that have occurred.
These two scores, evaluated at different thresholds, can be used to assess the quality of the model and adjust the threshold.
The balanced point (where Precision and Recall reach their highest values and the curves cross) is around 80 % for the Paper model and around 90 % for the Text model, which is another indicator that the Text model differentiates much better.

```{r agreementpapertextprecrec, echo=FALSE, message=FALSE, warnings=FALSE, results='asis', fig.cap="Precision and Recall for the Paper model and the Text model at varying theresholds; only for data used to estimate the model parameters and grouped by their actual outcome"}
.p <- predict(AgreementPaperModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementPaperModelPR <- lapply(.ts, .precrec)
.PrecisionPaper <- unlist(.AgreementPaperModelPR)[seq(1,201,2)]
.RecallPaper <- unlist(.AgreementPaperModelPR)[seq(2,202,2)]

.p <- predict(AgreementTextModel, type="response")
.precrec <- function(t) PrecisionRecall(.p, .d, t)
.AgreementTextModelPR <- lapply(.ts, .precrec)
.PrecisionText <- unlist(.AgreementTextModelPR)[seq(1,201,2)]
.RecallText <- unlist(.AgreementTextModelPR)[seq(2,202,2)]

par(mfrow=c(1,2))
plot(.PrecisionPaper, type="l", lwd=2, bty="n",
     main="Paper model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col=2
     )
lines(.RecallPaper, col=3, lwd=2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c(2,3), lwd=2, bty="n")
plot(.PrecisionText, type="l", lwd=2, bty="n",
     main="Text model",
     xlab="Probability threshold for prediction",
     ylab="Precision and Recall"
     ,ylim=c(0,1)
     ,xaxt="n", yaxt="n"
     ,col=2
     )
lines(.RecallText, col=3, lwd=2)
axis(1, at=seq(0,100,10),labels=seq(0,1,0.1))
axis(2, at=seq(0,1,0.1), labels=seq(0,100,10))
legend("bottomleft", legend=c("Precision", "Recall"), col=c(2,3), lwd=2, bty="n")
par(mfrow=c(1,1))
```


