---
title: "Underdetermination and statistical modelling in diachronic corpus linguistics: a reanalysis of Walkden et al. (2023)"
author: "Roland Schäfer (Friedrich-Schiller-Universität Jena)"
date: "Draft 8 of 31 May 2026"
mainfont: Libertinus Serif
sansfont: Libertinus Sans 
monofont: Inconsolata
mathfont: Libertinus Math
fontsize: 12pt
geometry: "left=3cm,right=3cm,top=2.5cm,bottom=2.5cm"
csl: unified-style-sheet-for-linguistics.csl
output:
  bookdown::pdf_document2: 
    latex_engine: xelatex
    number_sections: true
    keep_tex: yes
    toc: false
    extra_dependencies: ["float"]
  bookdown::word_document2:
    number_sections: true
  bookdown::html_document2: 
    number_sections: true
    self_contained: false
bibliography: 
  - Hypotheses.bib
  - rs.bib
editor_options: 
  markdown: 
    wrap: sentence
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = "")
set.seed(911)

library(lme4)
library(performance)
library(MuMIn)
library(vioplot)
library(LaplacesDemon)
library(knitr)
library(kableExtra)
library(colorspace)

the.lwd <- 3

load("Subject.RData")
load("Agreement.RData")

# Data reorganisation/renaming or prettier printing.
# Study 1
AgreementData$Dialect <- factor(AgreementData$Dialect, levels = c("Southern", "EM", "WM", "Northern"))
levels(AgreementData$Dialect) <- c("South", "WestMidlands", "EastMidlands", "North")
AgreementAggregateData$Dialect <- factor(AgreementAggregateData$Dialect, levels = c("Southern", "EM", "WM", "Northern"))
levels(AgreementAggregateData$Dialect) <- c("South", "EastMidlands", "WestMidlands", "North")
AgreementData$POS <- as.factor(AgreementData$POS)

# Study 2
colnames(SubjectData)[3] <- "Subject"
colnames(SubjectData)[7] <- "Text"
colnames(SubjectData)[10] <- "Macroregion"
SubjectData$Subject <- ifelse(SubjectData$Subject == "NULL", 0, 1)
levels(SubjectData$Genre) <- c("Nonliterary", "Literary", "Supplemental")

```

```{r, child=c('functions.Rmd')}
```

> *Abstract* I reanalyse the corpus data of @WalkdenEa2023, which uses three historical corpus studies to assess Trudgill's theory of sociolinguistic typology [@Trudgill2011], according to which short-term adult-driven language contact leads to simplification effects in language. I argue, on methodological and epistemological grounds, that Studies 1 and 2 provide no evidence for the hypotheses they are taken to support. The problem is twofold. At the level of inference, the paper reads low p-values from Null-Hypothesis Significance Testing as corroboration of a large, loosely specified network of hypotheses, when such a test has almost no capacity to expose those hypotheses as false if they were indeed false. At the level of the data, both studies mishandle aggregation: Study 1 collapses thousands of binary outcomes into a few dozen equally weighted per-text difference scores, and Study 2 uses an analogous aggregate as a regressor. Refitting without these errors, I find that little survives. In Study 1, once texts enter as a random effect, between-text variation swamps dialect, and the one generalisation that holds rests essentially on a single text. In Study 2, the reported model is all but indistinguishable from a model with only a random effect for individual texts and no other explanatory variables at all. I close with three suggested remedies: (i) disciplined, severity-aware inference paired with more numerically committed theories, (ii) open data as a routine part of peer review, and (iii) support for publishing negative findings. More broadly, however, historical-corpus data of this kind might rarely support a confirmatory framework, and the deeper error is forcing exploratory-grade data through a confirmatory machinery they were never meant to pass through.

```{r, child=c('1.philsci_rewrite.Rmd')}
```

```{r, child=c('3.models_rewrite.Rmd')}
```

# Data and inferences in @WalkdenEa2023 {#walkden}

```{r, child=c('4.study1_rewrite.Rmd')}
```

```{r, child=c('5.study2_rewrite.Rmd')}
```

```{r, child=c('6.gooduse_rewrite.Rmd')}
```

# Acknowledgments

I thank Felix Bildhauer (Leibniz-IDS Mannheim), Felix Golcher (Humboldt-Universität zu Berlin), Gert Webelhuth (Goethe-Universität Frankfurt) and the participants of a colloquium talk held at Goethe-Universität Frankfurt for their feedback and advice.
Remaining errors are exclusively my own responsibility.

# References
